\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming (CSEG1044)\\Unit 04 -- Lecture 01 Notes\\Generics Fundamentals (Generic Classes)}
\begin{document}
\maketitle
\tableofcontents

\section{Lecture Snapshot}
\begin{itemize}
  \item \textbf{Unit focus:} Generics, Lambdas, Swing, and JDBC
  \item \textbf{Main new ideas:} Why this topic exists, Attempt 1: separate classes for every type, Attempt 2: use one class with \texttt{Object}
  \item \textbf{Course thread:} Use Generics Fundamentals (Generic Classes) to move Campus Resource Manager toward a real desktop application with UI, callbacks, and persistence.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Lecture Overview}
Students usually find generics difficult for the wrong reason.
The syntax looks unusual, so they try to memorize angle brackets instead of understanding the design problem.

This lecture starts from the \textbf{problem} first:
\begin{enumerate}
  \item repeated classes for similar logic,
  \item a tempting but unsafe \texttt{Object}-based shortcut,
  \item a generic class that gives both reuse and type safety.
\end{enumerate}

Once that progression is clear, the syntax becomes much easier to read.
\notesources{\OOPSUnitIVSources}

\section{Prerequisite Bridge}
Before reading the new material in depth, do a fast recall of the older ideas listed below.
\begin{itemize}
  \item \textbf{Classes and methods}: A generic class is still a class; generics only add type parameters to familiar structures. Main reference: \texttt{Unit 02 / Lecture 01 slides/notes}.
  \item \textbf{Type relationships}: Generic type safety is easier to read once basic inheritance intuition is already clear. Main reference: \texttt{Unit 02 / Lecture 04 slides/notes}.
\end{itemize}
This lecture only revisits those ideas briefly so that the main explanation can stay focused on the new concept sequence.
\notesources{\OOPSUnitIVSources}

\section{Learning Outcomes}
After this lecture, you should be able to:
\begin{itemize}
  \item Explain why generic classes are better than duplicated classes or \texttt{Object}-only designs.
  \item Read and write a generic class such as \texttt{ResourceHolder<T>}.
  \item Distinguish \textbf{type parameter}, \textbf{type argument}, and \textbf{parameterized type}.
\end{itemize}

\section{Suggested Reading Order}
\begin{enumerate}
  \item Read the \textbf{Prerequisite Bridge} first and confirm each recalled idea.
  \item Study the central explanation in this order: \textit{Why this topic exists} $\rightarrow$ \textit{Attempt 1: separate classes for every type} $\rightarrow$ \textit{Attempt 2: use one class with \texttt{Object}} $\rightarrow$ \textit{Why \texttt{Object} is unsafe} $\rightarrow$ \textit{Attempt 3: use a generic class}.
  \item Trace the worked example line by line before checking short answers.
  \item Attempt the practice section without looking at the solution immediately.
  \item Finish with the exit questions and further-study suggestions to confirm retention.
\end{enumerate}
\notesources{\OOPSUnitIVSources}

\section{Key Terms}
\begin{itemize}
  \item \textbf{Generic class}: a class parameterized by one or more types
  \item \textbf{Type parameter}: the placeholder in the class declaration, such as \texttt{T}
  \item \textbf{Type argument}: the real type supplied later, such as \texttt{Book}
  \item \textbf{Parameterized type}: the typed form of a generic class, such as \texttt{ResourceHolder<Book>}
  \item \textbf{Raw type}: the generic class name used without type arguments, for example \texttt{ResourceHolder}
  \item \textbf{Wrapper class}: the object form of a primitive, for example \texttt{Integer} for \texttt{int}
\end{itemize}

\section{Detailed Notes}
\subsection{Why this topic exists}
Imagine that our project stores different campus resources:
\begin{itemize}
  \item books,
  \item laptops,
  \item lab keys,
  \item maybe later even ID cards or equipment requests.
\end{itemize}

Now suppose we want a simple class whose job is only this:
\begin{quote}
store one item and return one item
\end{quote}

The behavior is always the same.
Only the data type changes.

\subsection{Attempt 1: separate classes for every type}
A beginner may write:
\begin{itemize}
  \item \texttt{BookHolder}
  \item \texttt{LaptopHolder}
  \item \texttt{LabKeyHolder}
\end{itemize}

Each class has the same basic methods:
\begin{itemize}
  \item store an item,
  \item return the stored item.
\end{itemize}

\textbf{What is good here?}
\begin{itemize}
  \item The code is type-safe.
  \item A \texttt{BookHolder} clearly stores books only.
\end{itemize}

\textbf{What is the problem?}
\begin{itemize}
  \item The class logic is repeated.
  \item Every new data type requires another almost identical class.
  \item Maintenance becomes wasteful.
\end{itemize}

So separate classes give safety, but they do not scale well.

\subsection{Attempt 2: use one class with \texttt{Object}}
To remove duplication, many students try this design:
\begin{lstlisting}
class ResourceHolder {
    private Object item;

    void set(Object item) { this.item = item; }
    Object get() { return item; }
}
\end{lstlisting}

\textbf{Why this seems attractive:}
\begin{itemize}
  \item there is now only one class,
  \item \texttt{Object} can refer to many kinds of values,
  \item code duplication is reduced.
\end{itemize}

\textbf{But what do we lose?}
\begin{itemize}
  \item the class no longer knows what it is really storing,
  \item callers must cast the result,
  \item wrong casts are detected late at runtime.
\end{itemize}

\subsection{Why \texttt{Object} is unsafe}
Consider this example:
\begin{lstlisting}
ResourceHolder holder = new ResourceHolder();
holder.set(new Laptop("L-14"));

Book b = (Book) holder.get();
\end{lstlisting}

The compiler allows the cast syntax.
But the stored object is a \texttt{Laptop}, not a \texttt{Book}.

So the failure moves to runtime as \texttt{ClassCastException}.

\textbf{Important teaching point:}
\begin{quote}
\texttt{Object} reduces duplication, but it also removes useful type information.
\end{quote}

\subsection{Attempt 3: use a generic class}
This is exactly where generics help.

\begin{lstlisting}
class ResourceHolder<T> {
    private T item;

    void set(T item) { this.item = item; }
    T get() { return item; }
}
\end{lstlisting}

Now we still have:
\begin{itemize}
  \item one reusable class,
\end{itemize}
but we also get:
\begin{itemize}
  \item compile-time type checking.
\end{itemize}

\subsection{How to read the syntax correctly}
\paragraph{Step 1: class declaration.}
\texttt{class ResourceHolder<T>} means:
\begin{quote}
define a class that works for some type to be chosen later
\end{quote}

\paragraph{Step 2: type parameter.}
The symbol \texttt{T} is the \textbf{type parameter}.
It is only a placeholder name.
It does not mean a concrete class by itself.

\paragraph{Step 3: parameterized use.}
When we write:
\begin{lstlisting}
ResourceHolder<Book> bookHolder;
\end{lstlisting}
then:
\begin{itemize}
  \item \texttt{Book} is the \textbf{type argument},
  \item \texttt{ResourceHolder<Book>} is the \textbf{parameterized type}.
\end{itemize}

\paragraph{Step 4: mental model.}
You can read
\begin{lstlisting}
ResourceHolder<Book>
\end{lstlisting}
as:
\begin{quote}
a ResourceHolder whose item type is Book
\end{quote}

\subsection{Safe usage}
\begin{lstlisting}
ResourceHolder<Book> bookHolder = new ResourceHolder<>();
bookHolder.set(new Book("Clean Code"));
Book b = bookHolder.get();
\end{lstlisting}

Notice what changed compared to the \texttt{Object} version:
\begin{itemize}
  \item no cast is required when calling \texttt{get()},
  \item the compiler knows the result should be a \texttt{Book},
  \item a wrong assignment can be caught earlier.
\end{itemize}

\subsection{Why generic classes improve teaching clarity}
Generic classes do not only improve safety.
They also improve readability.

Compare these two declarations:
\begin{lstlisting}
ResourceHolder rawHolder;
ResourceHolder<Book> bookHolder;
\end{lstlisting}

The second line communicates design intent immediately.
That matters in real projects, especially when classes are used by other developers.

\subsection{Common naming conventions}
Java commonly uses short uppercase letters for type parameters:
\begin{itemize}
  \item \texttt{T} = Type
  \item \texttt{E} = Element
  \item \texttt{K} = Key
  \item \texttt{V} = Value
\end{itemize}

These names are conventions, not magic keywords.

\subsection{Primitive types do not work directly}
This is a frequent beginner confusion.

\begin{lstlisting}
ResourceHolder<int> marks;      // invalid
ResourceHolder<Integer> marks;  // correct
\end{lstlisting}

Generics require reference types.
Primitive values must be wrapped using classes such as:
\begin{itemize}
  \item \texttt{Integer}
  \item \texttt{Double}
  \item \texttt{Character}
  \item \texttt{Boolean}
\end{itemize}

\subsection{Raw types and why they are dangerous}
Raw type means using the class name without a type argument:
\begin{lstlisting}
ResourceHolder raw = new ResourceHolder(new LabKey("LAB-2"));
\end{lstlisting}

\textbf{Why Java still allows raw types:}
\begin{itemize}
  \item backward compatibility with old pre-generics code
\end{itemize}

\textbf{Why you should usually avoid them:}
\begin{itemize}
  \item they disable the main safety advantage of generics,
  \item they cause warnings,
  \item they make later casts risky.
\end{itemize}

\subsection{A useful intuition about compile-time vs runtime}
Generics in Java are mainly a \textbf{compile-time safety feature}.
For this lecture, the key takeaway is simple:
\begin{itemize}
  \item without generics, many mistakes survive until runtime,
  \item with generics, many of those mistakes are caught earlier.
\end{itemize}

That is the pedagogically important point for now.
More advanced runtime details such as type erasure will be revisited later.
\notesources{\OOPSUnitIVSources}

\section{Running Project Connection}
Use Generics Fundamentals (Generic Classes) to move Campus Resource Manager toward a real desktop application with UI, callbacks, and persistence.
\begin{itemize}
  \item Use Campus Resource Manager as the running case study, or map the same idea to your own project domain if you are using another example.
  \item Ask where today's idea should live: model, service, DAO, UI, or team workflow.
  \item Use the lecture example as a smaller version of the same design choice you will make in the capstone.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Common Mistakes and Confusions}
\begin{itemize}
  \item Treating generic type parameters as if they were inheritance declarations.
  \item Falling back to raw types and losing type safety.
  \item Expecting primitive types to work directly as generic arguments.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Worked Example: campus resources}
\subsection*{Domain classes}
\begin{lstlisting}
class Book {
    String title;
    Book(String title) { this.title = title; }
}

class LabKey {
    String roomCode;
    LabKey(String roomCode) { this.roomCode = roomCode; }
}
\end{lstlisting}

\subsection*{Generic class}
\begin{lstlisting}
class ResourceHolder<T> {
    private final T item;

    ResourceHolder(T item) { this.item = item; }
    T get() { return item; }
}
\end{lstlisting}

\subsection*{Safe usage}
\begin{lstlisting}
public class DemoSafe {
    public static void main(String[] args) {
        ResourceHolder<Book> b =
            new ResourceHolder<>(new Book("Clean Code"));
        ResourceHolder<LabKey> k =
            new ResourceHolder<>(new LabKey("LAB-2"));

        Book book = b.get();
        LabKey key = k.get();
        System.out.println(book.title + " / " + key.roomCode);
    }
}
\end{lstlisting}

\subsection*{Unsafe raw-type usage}
\begin{lstlisting}
public class DemoRaw {
    public static void main(String[] args) {
        ResourceHolder raw =
            new ResourceHolder(new LabKey("LAB-2"));

        Book wrong = (Book) raw.get();
        System.out.println(wrong.title);
    }
}
\end{lstlisting}

\paragraph{What this example teaches.}
\begin{itemize}
  \item The generic version keeps one reusable class.
  \item Different type arguments create different typed usages.
  \item The raw version removes the generic benefit and reintroduces risk.
\end{itemize}

\section{Demo / Observation Task}
Use this block as a short reproducible demo or observation task.
\begin{itemize}
  \item Reproduce the lecture's main example around \textit{Why this topic exists} once without looking at the solution first.
  \item Change one input, rule, or design choice in Campus Resource Manager and predict the result before checking it.
  \item Record one success case, one edge case, and one failure case so the idea becomes observable instead of purely theoretical.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Practice Check (with brief solutions)}
\begin{enumerate}
  \item Why is a single \texttt{Object}-based holder better than many repeated holder classes, but still not the best final solution?\par
  \textbf{Answer:} It removes duplication, but it loses type information and forces unsafe casts.

  \item What is the difference between a type parameter and a type argument?\par
  \textbf{Answer:} The type parameter is the placeholder in the class declaration, such as \texttt{T}; the type argument is the real type supplied later, such as \texttt{Book}.

  \item Why do we write \texttt{ResourceHolder<Integer>} instead of \texttt{ResourceHolder<int>}?\par
  \textbf{Answer:} Generic type arguments must be reference types, so we use the wrapper class \texttt{Integer} instead of the primitive \texttt{int}.
\end{enumerate}

\section{Self-Study Practice (no solutions)}
\begin{enumerate}
  \item Create a generic class \texttt{Pair<K,V>} that stores two related values.
  \item Write a generic \texttt{TicketHolder<T>} and instantiate it with two different user-defined classes.
  \item Find one old example that uses raw collections and rewrite it with parameterized types.
  \item Explain in your own words why ``one class with \texttt{Object}'' is not the same as a generic class.
\end{enumerate}

\section{Further Study}
\begin{itemize}
  \item Revisit the prerequisite references if any recall bullet still feels weak.
  \item Rework the lecture example with one changed assumption, input, or design choice.
  \item Read the textbook and documentation references listed in the source line for a second explanation of the same topic.
  \item Apply the same idea once inside Campus Resource Manager so that the concept moves from theory to design practice.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Exit Questions (with sample answers)}
\textbf{Q1.} What design problem leads programmers toward generic classes?\par
\textbf{Sample answer:} The same class logic is often needed for many data types. Writing separate classes for each type causes duplication, while using \texttt{Object} removes safety. Generic classes solve both issues together.

\vspace{0.6em}
\textbf{Q2.} What does \texttt{ResourceHolder<Book>} mean?\par
\textbf{Sample answer:} It means a typed version of \texttt{ResourceHolder} whose stored item is expected to be a \texttt{Book}. Here, \texttt{Book} is the type argument.

\vspace{0.6em}
\textbf{Q3.} Why should modern Java code avoid raw types where possible?\par
\textbf{Sample answer:} Raw types weaken compile-time checks, create warnings, and make unsafe casts possible again. They remove the main advantage that generics were introduced to provide.

\end{document}
