\documentclass{beamer}

\input{../../../_shared/beamer-common.tex}

\title[OOPS with Java]{Object Oriented Programming (CSEG1044)}
\subtitle{Unit 03 -- Lecture 03: Multithreading + Synchronization}
\author{Tofik Ali}
\institute{School of Computer Science, UPES Dehradun}
\date{\today}

\begin{document}

\begin{frame}[fragile]
  \titlepage
  \vspace{0.5em}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Quick Links}
  \centering
  \hyperlink{sec:overview}{\beamerbutton{Overview}}\hspace{0.6em}
  \hyperlink{sec:concepts}{\beamerbutton{Key Topics}}\hspace{0.6em}
  \hyperlink{sec:demo}{\beamerbutton{Demo}}\hspace{0.6em}
  \hyperlink{sec:summary}{\beamerbutton{Summary}}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Agenda}
  \tableofcontents
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\section{Overview}
\label{sec:overview}

\begin{frame}{Learning Outcomes}
  \begin{itemize}[<+->]
    \item Create threads using \texttt{Thread} and \texttt{Runnable}.
    \item Describe thread lifecycle and understand priority limitations.
    \item Identify race conditions and fix them with \texttt{synchronized}.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\section{Key Topics}
\label{sec:concepts}

\begin{frame}{Unit 03: Nested Classes, Exceptions, Multithreading \& IO Streams}
  \begin{itemize}[<+->]
    \item Lecture focus: Multithreading + Synchronization
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Today: Roadmap}
  \begin{itemize}[<+->]
    \item Concurrency motivation; `Thread` class and `Runnable` interface
    \item Thread lifecycle and priorities (what they mean / limitations)
    \item Race conditions and synchronization (`synchronized`, basic patterns)
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Concurrency vs Parallelism}
  \begin{itemize}[<+->]
    \item \textbf{Concurrency}: tasks overlap in time (good for responsiveness).
    \item \textbf{Parallelism}: tasks run simultaneously on multiple cores (good for speed).
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Creating threads (two ways)}
  \begin{itemize}[<+->]
    \item Extend \texttt{Thread} and override \texttt{run()}.
    \item Implement \texttt{Runnable} and pass it to \texttt{new Thread(r)}.
    \item \textbf{Important:} call \texttt{start()} to create a new thread.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{\texttt{start()} vs \texttt{run()} (very important)}
\begin{lstlisting}
Runnable task = new Runnable() {
    public void run() {
        System.out.println(Thread.currentThread().getName());
    }
};

Thread t = new Thread(task, "worker");

t.run();   // runs on current thread (main)
t.start(); // runs on a new thread (worker)
\end{lstlisting}
  \begin{itemize}[<+->]
    \item \texttt{run()} is a normal method call (no new thread).
    \item \texttt{start()} asks JVM/OS to schedule a new thread.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Waiting for threads: \texttt{join()}}
\begin{lstlisting}
Runnable task = new Runnable() {
    public void run() { /* do work */ }
};

Thread t1 = new Thread(task);
Thread t2 = new Thread(task);
t1.start(); t2.start();

t1.join();  // wait for t1 to finish
t2.join();  // wait for t2 to finish
System.out.println("Both done");
\end{lstlisting}
  \begin{itemize}[<+->]
    \item Without \texttt{join()}, main may finish early.
    \item \texttt{join()} throws \texttt{InterruptedException}.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Lifecycle + priorities (practical)}
  \begin{itemize}[<+->]
    \item New $\rightarrow$ Runnable/Running $\rightarrow$ Waiting/Blocked $\rightarrow$ Terminated
    \item Waiting happens due to sleep, locks, I/O, etc.
    \item Priority is a \textbf{hint}, not a guarantee (depends on OS/JVM).
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Race condition (core problem)}
  \begin{itemize}[<+->]
    \item Multiple threads modify shared state at the same time.
    \item Result depends on timing/interleaving.
    \item Example: \texttt{value++} is not atomic (read-modify-write).
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Synchronization with \texttt{synchronized}}
  \begin{itemize}[<+->]
    \item Protect a critical section using a lock/monitor.
    \item Only one thread at a time can execute synchronized code on the same lock.
    \item Keep synchronized sections small; avoid deadlocks.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}


\section{Demo / Example}
\label{sec:demo}

\begin{frame}[fragile]{Example: synchronized counter}
\begin{lstlisting}
class SafeCounter {
    private int value = 0;
    synchronized void inc() { value++; }
    synchronized int get() { return value; }
}
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Thread demo (two threads) (1/2)}
\begin{lstlisting}
public class Demo {
    public static void main(String[] args) throws InterruptedException {
        SafeCounter c = new SafeCounter();

        Runnable task = new Runnable() {
            public void run() {
                for (int i = 0; i < 100000; i++) c.inc();
            }
        };
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Thread demo (two threads) (2/2)}
\begin{lstlisting}
// continued...
        Thread t1 = new Thread(task);
        Thread t2 = new Thread(task);
        t1.start();
        t2.start();
        t1.join(); t2.join();
        System.out.println(c.get()); // expected 200000
    }
}
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Mini Exercises (2--3 mins)}
  \begin{enumerate}[<+->]
    \item Difference: \texttt{start()} vs \texttt{run()}?
    \item Why is \texttt{value++} unsafe in multiple threads?
    \item What does \texttt{synchronized} guarantee (in one sentence)?
  \end{enumerate}
  \slidesources{\OOPSUnitIIISources}
\end{frame}


\section{Summary}
\label{sec:summary}

\begin{frame}{Key Takeaways}
  \begin{itemize}[<+->]
    \item Threads enable concurrency for responsiveness and throughput.
    \item Race conditions occur with shared mutable state.
    \item \texttt{synchronized} protects critical sections using a lock.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Checkpoint Questions}
  \begin{enumerate}[<+->]
    \item What is a race condition?
    \item Why is thread priority not a guarantee?
    \item How does \texttt{synchronized} help?
  \end{enumerate}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Next Lecture Preview}
  \textbf{Next:} Java I/O Streams + FileReader/FileWriter
  \vspace{0.6em}
  \begin{itemize}[<+->]
    \item We will read/write files and handle errors safely with exceptions.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\end{document}
