\documentclass{beamer}

\input{../../../_shared/beamer-common.tex}

\title[OOPS with Java]{Object Oriented Programming (CSEG1044)}
\subtitle{Unit 03 -- Lecture 01: Nested Classes (Types + Use-cases)}
\begin{document}

\begin{frame}[fragile]
  \titlepage
  \vspace{0.5em}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Quick Links}
  \centering
  \hyperlink{sec:overview}{\beamerbutton{Overview}}\hspace{0.6em}
  \hyperlink{sec:concepts}{\beamerbutton{Key Topics}}\hspace{0.6em}
  \hyperlink{sec:demo}{\beamerbutton{Demo}}\hspace{0.6em}
  \hyperlink{sec:summary}{\beamerbutton{Summary}}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Agenda}
  \tableofcontents
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\section{Overview}
\label{sec:overview}

\begin{frame}{Lecture Snapshot}
\begin{itemize}[<+->]
  \item \textbf{Unit focus:} Nested Classes, Exceptions, Multithreading, and I/O
  \item \textbf{Course thread:} Use Nested Classes (Types + Use-cases) to make Campus Resource Manager safer, more modular, and easier to recover when something goes wrong.
\end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Recall Before You Start}
\begin{itemize}[<+->]
  \item \textbf{Class structure}: Nested classes still depend on the class-object model from Unit 02. Main reference: \texttt{Unit 02 / Lecture 01 slides/notes}.
  \item \textbf{Interfaces and callbacks}: Anonymous and local classes make more sense once contract-based behavior is familiar. Main reference: \texttt{Unit 02 / Lecture 05 slides/notes}.
\end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}


\begin{frame}{Learning Outcomes}
  \begin{itemize}[<+->]
    \item Explain why nested classes improve encapsulation in some designs.
    \item Differentiate static nested, inner, local, and anonymous classes.
    \item Recognize use cases: helpers and callbacks/listeners (Swing later).
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Key Terms to Know}
\small
\begin{columns}[t]
\column{0.48\textwidth}
\begin{itemize}
  \item \textbf{Static nested class}: nested class with \texttt{static}; no implicit outer instance.
  \item \textbf{Inner class}: non-static member class; has implicit reference to outer object.
  \item \textbf{Local class}: class declared inside a method.
  \item \textbf{Anonymous class}: unnamed local class used inline to implement an interface/extend a class.
\end{itemize}
\column{0.48\textwidth}
\begin{itemize}
  \item \textbf{Callback}: passing code/behavior to run later (often via interfaces).
  \item \textbf{Effectively final}: a local variable that is not reassigned after initialization.
  \item \textbf{Outer instance}: the object of the outer class that an inner class ``belongs to''.
\end{itemize}
\end{columns}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Study Flow for This Lecture}
\begin{enumerate}[<+->]
  \item Refresh the prerequisite bridge before reading new ideas in isolation.
  \item Study the main build in this order: \textit{Nested classes at a glance} $\rightarrow$ \textit{Motivation: scope and encapsulation} $\rightarrow$ \textit{Types of nested classes (with intuition)} $\rightarrow$ \textit{Access rules (important and often surprising)}.
  \item Trace the worked example and connect each line back to one concept frame.
  \item Attempt the mini exercises before moving to the recap and exit check.
\end{enumerate}
  \slidesources{\OOPSUnitIIISources}
\end{frame}


\section{Key Topics}
\label{sec:concepts}

\begin{frame}{Unit 03: Nested Classes, Exceptions, Multithreading \& IO Streams}
  \begin{itemize}[<+->]
    \item Lecture focus: Nested Classes (Types + Use-cases)
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Today: Roadmap}
  \begin{itemize}[<+->]
    \item Nested class motivation: scope and encapsulation
    \item Static nested, inner, local, and anonymous classes
    \item Use-cases: helpers, callbacks/listeners (bridge to Swing later)
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Why nested classes?}
  \begin{itemize}[<+->]
    \item Keep helper types close to the code that uses them.
    \item Reduce public API surface (better encapsulation).
    \item Common for callbacks/listeners.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Types (quick)}
  \begin{itemize}[<+->]
    \item \textbf{Static nested}: no outer instance reference.
    \item \textbf{Inner}: has implicit outer instance reference.
    \item \textbf{Local}: declared inside a method; limited scope.
    \item \textbf{Anonymous}: unnamed; used inline for one-time behavior.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Static nested vs inner (key difference)}
  \begin{itemize}[<+->]
    \item \textbf{Static nested}: behaves like a normal class scoped inside outer; no outer object needed.
    \item \textbf{Inner}: every instance is tied to an outer object (has an implicit outer reference).
  \end{itemize}
  \vspace{0.6em}
  \textbf{Rule:} if you do not need outer state, prefer \texttt{static}.
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Example: static nested helper (validator)}
\begin{lstlisting}
class BankAccount {
  static class Validator {
    static void requirePositive(double amt) {
      if (amt <= 0) throw new IllegalArgumentException("amt");
    }
  }

  void deposit(double amt) {
    Validator.requirePositive(amt);
    // ... update balance
  }
}
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Example: local class (inside a method)}
\begin{lstlisting}
public void demoLocal() {
  final int times = 2; // explicitly final

  class Printer {
    void say(String msg) {
      for (int i = 0; i < times; i++)
        System.out.println(msg);
    }
  }

  new Printer().say("Hi");
}
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Use cases}
  \begin{itemize}[<+->]
    \item Helpers: validators, builders, entries/tuples
    \item Iterators and small adapters
    \item Callbacks/listeners (Swing event handling later)
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Checkpoint and Reflection}
\begin{block}{Reflection prompt}
Where would \textit{Nested Classes (Types + Use-cases)} matter most in Campus Resource Manager, and which design choice would you justify using this lecture's ideas?
\end{block}
\begin{block}{Quick self-check}
Which part needs one more example before you move on: \textit{Nested classes at a glance}, \textit{Motivation: scope and encapsulation}, the worked example, or the recap?
\end{block}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\section{Demo / Example}
\label{sec:demo}

\begin{frame}[fragile]{Example: inner + anonymous (small) (1/2)}
\begin{lstlisting}
interface Greeter { void greet(String name); }

public class Outer {
    private final String prefix = "Hello";

    class InnerGreeter implements Greeter {
        public void greet(String name) {
            System.out.println(prefix + ", " + name);
        }
    }
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Example: inner + anonymous (small) (2/2)}
\begin{lstlisting}
// continued...
    public void demo() {
        Greeter a = new InnerGreeter();
        Greeter b = new Greeter() { // anonymous
            public void greet(String name) {
                System.out.println(prefix + " (anon), " + name);
            }
        };
        a.greet("UPES");
        b.greet("UPES");
    }
}
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Mini Exercises (2--3 mins)}
  \begin{enumerate}[<+->]
    \item Which type has an implicit outer reference?
    \item Which type is commonly used for callbacks?
    \item What does ``effectively final'' mean for local classes?
  \end{enumerate}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Watch-outs}
\begin{itemize}[<+->]
  \item Confusing static nested classes with true inner classes.
  \item Overusing anonymous classes when the logic is too large to read.
  \item Scattering tiny helper types across many files without need.
\end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\section{Summary}
\label{sec:summary}

\begin{frame}{Key Takeaways}
  \begin{itemize}[<+->]
    \item Nested classes improve encapsulation and keep helpers local.
    \item Know the 4 kinds: static nested, inner, local, anonymous.
    \item Anonymous classes are common for callback-style code (Swing).
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Checkpoint Questions}
  \begin{enumerate}[<+->]
    \item Difference: static nested vs inner class?
    \item Why are anonymous classes used in event handling?
    \item Where can a local class be used?
  \end{enumerate}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Next Study Steps}
\begin{itemize}[<+->]
  \item Revisit the recall references if any older idea still feels weak.
  \item Rework the lecture example without looking at the notes first.
  \item Attempt the self-study practice and then answer the exit questions in full sentences.
  \item Apply the idea once inside Campus Resource Manager to make the concept stick.
\end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Next Lecture Preview}
  \textbf{Next:} Exception Handling (Core)
  \vspace{0.6em}
  \begin{itemize}[<+->]
    \item We will handle errors safely using \texttt{try/catch/finally} and custom exceptions.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\end{document}
