\documentclass{beamer}

\input{../../../_shared/beamer-common.tex}

\title[OOPS with Java]{Object Oriented Programming (CSEG1044)}
\subtitle{Unit 06 -- Lecture 06: Testing + Build/Artifacts + Final Demo}
\author{Tofik Ali}
\institute{School of Computer Science, UPES Dehradun}
\date{\today}

\begin{document}

\begin{frame}[fragile]
  \titlepage
  \vspace{0.5em}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Quick Links}
  \centering
  \hyperlink{sec:overview}{\beamerbutton{Overview}}\hspace{0.6em}
  \hyperlink{sec:concepts}{\beamerbutton{Key Topics}}\hspace{0.6em}
  \hyperlink{sec:demo}{\beamerbutton{Demo}}\hspace{0.6em}
  \hyperlink{sec:summary}{\beamerbutton{Summary}}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Agenda}
  \tableofcontents
  \slidesources{\OOPSUnitVISources}
\end{frame}

\section{Overview}
\label{sec:overview}

\begin{frame}{Learning Outcomes}
  \begin{itemize}[<+->]
    \item Write JUnit tests for service/business rules.
    \item Run an integration checklist for DAO + DB.
    \item Package a runnable artifact and prepare a final demo (Deliverable-5).
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Goal: demo-ready project}
  \begin{itemize}[<+->]
    \item Core flows are reliable (not ``works only sometimes'').
    \item Errors are handled gracefully (validation + messages).
    \item Build produces a runnable artifact (Jar/zip).
    \item You have a short demo script and clear evidence of teamwork.
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Deliverable-5 (final submission)}
  \begin{itemize}[<+->]
    \item Runnable artifact + run instructions.
    \item \texttt{schema.sql} + DB setup notes.
    \item Unit tests (JUnit) for service rules.
    \item Short report + README + team contributions.
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\section{Key Topics}
\label{sec:concepts}

\begin{frame}{Unit 06: Capstone Project}
  \begin{itemize}[<+->]
    \item Lecture focus: Testing + Build/Artifacts + Final Demo
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Today: Roadmap}
  \begin{itemize}[<+->]
    \item Testing strategy: what to unit test vs integrate
    \item JUnit 5 essentials (assertions + test structure)
    \item Build + packaging + demo checklist (Deliverable-5)
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Testing strategy (capstone)}
  \begin{itemize}[<+->]
    \item \textbf{Unit tests (fast):} service validation + business rules.
    \item \textbf{Integration tests:} DAO + DB schema + mapping.
    \item \textbf{Manual demo:} Swing flows (minimal but polished).
  \end{itemize}
  \textbf{Rule:} if a bug can be caught by a unit test, write the unit test.
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Test pyramid (in this course)}
  \centering
  \begin{tikzpicture}[every node/.style={font=\small}]
    \draw[fill=blue!10, draw=black] (0,0) rectangle (5,0.9);
    \draw[fill=green!10, draw=black] (0.6,0.9) rectangle (4.4,1.8);
    \draw[fill=orange!10, draw=black] (1.2,1.8) rectangle (3.8,2.7);

    \node at (2.5,0.45) {\textbf{Unit tests (many)}};
    \node at (2.5,1.35) {\textbf{Integration tests (few)}};
    \node at (2.5,2.25) {\textbf{Manual UI demo (minimal)}};
  \end{tikzpicture}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{JUnit 5 essentials}
  \begin{itemize}[<+->]
    \item Use \texttt{@Test} methods; follow Arrange--Act--Assert.
    \item Assertions: \texttt{assertEquals}, \texttt{assertTrue}, \texttt{assertThrows}.
    \item Use \texttt{@BeforeEach} to set up a clean fixture.
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}[fragile]{Example: \texttt{assertThrows} for validation}
\begin{lstlisting}
@Test
void rejectsEndBeforeStart() {
    assertThrows(ValidationException.class,
        () -> service.createBooking(1, "A", 10, 10));
}
\end{lstlisting}
  \begin{itemize}[<+->]
    \item Good tests are small and check one behavior.
    \item Name tests like a sentence: \texttt{rejectsEndBeforeStart}.
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Design for testability: interface + fake}
  \centering
  \begin{tikzpicture}[node distance=1.3cm, every node/.style={font=\small}]
    \node[draw, rounded corners, fill=green!8, inner sep=6pt] (svc) {BookingService};
    \node[draw, rounded corners, fill=yellow!10, right=of svc, inner sep=6pt] (iface) {BookingRepository (interface)};
    \node[draw, rounded corners, fill=blue!8, above right=0.5cm and 0.8cm of iface, inner sep=6pt] (fake) {InMemoryRepo (tests)};
    \node[draw, rounded corners, fill=orange!10, below right=0.5cm and 0.8cm of iface, inner sep=6pt] (jdbc) {JdbcRepo (app)};
    \draw[-{Latex[length=3mm]}] (svc) -- (iface);
    \draw[-{Latex[length=3mm]}] (iface) -- (fake);
    \draw[-{Latex[length=3mm]}] (iface) -- (jdbc);
  \end{tikzpicture}

  \vspace{0.6em}
  \begin{itemize}[<+->]
    \item Tests run fast without DB.
    \item Production uses JDBC implementation.
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Integration checklist (DAO + DB)}
  \begin{enumerate}[<+->]
    \item Run \texttt{schema.sql} on a clean DB.
    \item Insert one row using DAO.
    \item List/find and verify mapping fields.
    \item Update and verify changes.
    \item Delete and verify removal.
  \end{enumerate}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Build + packaging (idea)}
  \begin{itemize}[<+->]
    \item Maven: \texttt{mvn test}, then \texttt{mvn package} (Jar in \texttt{target/})
    \item Gradle: \texttt{gradle test}, then \texttt{gradle build}
    \item Artifact should be runnable with clear instructions.
  \end{itemize}
  \textbf{Tip:} record your run steps in README (so TA can run it quickly).
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Release + demo preparation}
  \begin{itemize}[<+->]
    \item Freeze scope: choose what you will demo.
    \item Run tests + integration checklist.
    \item Prepare demo data (1--2 example resources, bookings).
    \item Create a release (tag) and attach the artifact.
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\section{Demo / Example}
\label{sec:demo}

\begin{frame}{Demo script (5--7 mins)}
  \begin{enumerate}[<+->]
    \item Start DB; confirm schema exists.
    \item Run the app; show one end-to-end flow (Add Resource $\rightarrow$ list).
    \item Show booking flow (create + list).
    \item Show validation (invalid booking shows a clean message).
    \item (Optional) Run tests: \texttt{mvn test}.
  \end{enumerate}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Mini Exercises (5 mins)}
  \begin{enumerate}[<+->]
    \item Write one unit test for a service rule in your project.
    \item What is one integration test you must do for DAO?
    \item What 3 steps should be in your README ``How to run''?
  \end{enumerate}
  \slidesources{\OOPSUnitVISources}
\end{frame}


\section{Summary}
\label{sec:summary}

\begin{frame}{Key Takeaways}
  \begin{itemize}[<+->]
    \item Test service rules with JUnit; keep tests fast and focused.
    \item Verify DAO with an integration checklist.
    \item Package a runnable artifact and rehearse a short demo script.
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Checkpoint Questions}
  \begin{enumerate}[<+->]
    \item What is the difference between unit and integration tests?
    \item Why does ``depend on an interface'' help testing?
    \item What must be included in a demo-ready release artifact?
  \end{enumerate}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\begin{frame}{Course Wrap-Up}
  \begin{itemize}[<+->]
    \item Use OOP principles to keep code clean and reusable.
    \item Use interfaces and layering to reduce coupling.
    \item Use testing + packaging to ship reliable software.
  \end{itemize}
  \slidesources{\OOPSUnitVISources}
\end{frame}

\end{document}
