\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming Lab (CSEG1044)\\Experiment-01\&02: Java Programming Environment + Input Handling (Student Handout --- Before Submission)}
\author{Tofik Ali}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

\section{About this handout}
This is a \textbf{pre-submission} student handout.
It provides \textbf{requirements} and \textbf{hints}, but not full solutions.
Teacher model solutions are kept privately.

\section{Experiment 01 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To understand the Java programming environment, program structure, compilation, and execution process.
  \item \textbf{Problem Statement:} Design and implement a basic Java program demonstrating class declaration, \texttt{main()} method, and standard output statements.
  \item \textbf{Expected Outcome:} Ability to write, compile, and execute a basic Java program.
\end{itemize}

\section{Experiment 02 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To understand different methods of input handling in Java.
  \item \textbf{Problem Statement:} Develop a Java program that accepts input using standard input methods and command line arguments.
  \item \textbf{Expected Outcome:} Ability to handle input from multiple sources.
\end{itemize}

\section{What you must do (minimum requirements)}

\subsection{Task A --- Java setup + first program (Exp-01)}
\begin{enumerate}[leftmargin=*]
  \item Verify installation (attach output screenshots):
  \begin{itemize}[leftmargin=*]
    \item \texttt{java -version}
    \item \texttt{javac -version}
  \end{itemize}

  \item Create \texttt{HelloUPES.java} that prints:
  \begin{itemize}[leftmargin=*]
    \item a welcome message,
    \item Java version using \texttt{System.getProperty("java.version")},
    \item number of command line arguments.
  \end{itemize}

  \item Compile and run:
  \begin{itemize}[leftmargin=*]
    \item \texttt{javac HelloUPES.java}
    \item \texttt{java HelloUPES}
    \item \texttt{java HelloUPES A B}
  \end{itemize}
\end{enumerate}

\subsection{Task B --- Keyboard input + command line args (Exp-02)}
\begin{enumerate}[leftmargin=*]
  \item Create \texttt{InputArgsDemo.java} that:
  \begin{itemize}[leftmargin=*]
    \item reads \texttt{bonusPercent} from command line arguments,
    \item reads \texttt{name} and \texttt{baseSalary} using \texttt{Scanner},
    \item computes \texttt{finalSalary = baseSalary + baseSalary * bonusPercent / 100},
    \item prints the final salary using formatted output.
  \end{itemize}

  \item Validations (mandatory):
  \begin{itemize}[leftmargin=*]
    \item check \texttt{args.length} before accessing \texttt{args[0]},
    \item handle invalid bonus argument (\texttt{NumberFormatException}),
    \item handle invalid salary input (\texttt{InputMismatchException}).
  \end{itemize}
\end{enumerate}

\section{Hints (read only if stuck)}
\begin{itemize}[leftmargin=*]
  \item Command line arguments are always strings; parse numbers using \texttt{Integer.parseInt} or \texttt{Double.parseDouble}.
  \item Close \texttt{Scanner} using \texttt{finally} or try-with-resources.
  \item If you mix \texttt{nextInt()/nextDouble()} with \texttt{nextLine()}, remember the newline issue.
\end{itemize}

\section{What to submit (MANDATORY)}
Submit a single zipped folder containing:
\begin{itemize}[leftmargin=*]
  \item source code: \texttt{HelloUPES.java}, \texttt{InputArgsDemo.java}
  \item output screenshots or copied output
  \item observations (5--10 lines): compile-time vs runtime; what happens for invalid inputs
\end{itemize}

\notesources{\OOPSUnitISources}
\end{document}
