\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming Lab (CSEG1044)\\Experiment-11\&12: Exception Handling + Multithreading (Student Handout --- Before Submission)}
\author{Tofik Ali}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

\section{About this handout}
This handout describes the required programs for exceptions and multithreading.
It provides requirements and common pitfalls, but not full solutions.

\section{Experiment 11 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To handle runtime errors using exception handling mechanisms.
  \item \textbf{Problem Statement:} Develop programs using try-catch-finally blocks.
  \item \textbf{Expected Outcome:} Students will write robust programs.
\end{itemize}

\section{Experiment 12 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To implement multithreading in Java.
  \item \textbf{Problem Statement:} Create threads using \texttt{Thread} class and \texttt{Runnable} interface.
  \item \textbf{Expected Outcome:} Students will understand concurrent programming.
\end{itemize}

\section{What you must do (minimum requirements)}

\subsection{Task A --- Exception handling (Exp-11)}
Create \texttt{ExceptionHandlingDemo.java}:
\begin{enumerate}[leftmargin=*]
  \item Read integers \texttt{a} and \texttt{b}, compute \texttt{a / b}.
  \item Handle:
  \begin{itemize}[leftmargin=*]
    \item divide-by-zero (\texttt{ArithmeticException})
    \item invalid numeric input (scanner mismatch or parse error)
  \end{itemize}
  \item Use \texttt{finally} to print \texttt{Done} (and close resources if any).
  \item Create a custom checked exception \texttt{InvalidMarksException extends Exception}.
  \item Write \texttt{validateMarks(int m) throws InvalidMarksException} and throw it for marks not in \texttt{0..100}.
  \item Catch and print a friendly message.
\end{enumerate}

\subsection{Task B --- Multithreading + synchronization (Exp-12)}
Create \texttt{ThreadSyncDemo.java}:
\begin{enumerate}[leftmargin=*]
  \item Create a shared \texttt{Counter} with:
  \begin{itemize}[leftmargin=*]
    \item \texttt{incUnsafe()} (no synchronization)
    \item \texttt{incSafe()} (use \texttt{synchronized})
  \end{itemize}
  \item Create two threads that increment the same counter many times (example: 100000 each).
  \item Use \texttt{start()} and \texttt{join()}.
  \item Print final counter values for unsafe and safe runs.
  \item Write a short explanation (2--4 lines): why unsafe output differs from expected.
\end{enumerate}

\section{Hints (read only if stuck)}
\begin{itemize}[leftmargin=*]
  \item \texttt{finally} executes whether an exception occurs or not.
  \item Checked exceptions must be handled or declared using \texttt{throws}.
  \item Calling \texttt{run()} does not start a new thread; call \texttt{start()}.
  \item \texttt{v++} is not atomic; \texttt{synchronized} protects the critical section.
\end{itemize}

\section{What to submit (MANDATORY)}
Submit a single zipped folder containing:
\begin{itemize}[leftmargin=*]
  \item \texttt{ExceptionHandlingDemo.java}
  \item \texttt{ThreadSyncDemo.java}
  \item sample output for both programs
  \item observations (5--10 lines): try/catch/finally order; what race condition means
\end{itemize}

\notesources{\OOPSUnitIIISources}
\end{document}
