\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming Lab (CSEG1044)\\Experiment-09\&10: Inheritance + Abstraction (Student Handout --- Before Submission)}
\author{Tofik Ali}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

\section{About this handout}
This handout describes what to implement for inheritance, overriding, abstract classes, and interfaces.
It intentionally avoids giving full code.

\section{Experiment 09 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To implement inheritance and runtime polymorphism.
  \item \textbf{Problem Statement:} Create base and derived classes demonstrating inheritance.
  \item \textbf{Expected Outcome:} Students will understand code reusability.
\end{itemize}

\section{Experiment 10 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To implement abstraction using abstract classes and interfaces.
  \item \textbf{Problem Statement:} Design programs using abstract classes and interfaces.
  \item \textbf{Expected Outcome:} Students will understand abstraction and multiple inheritance.
\end{itemize}

\section{What you must do (minimum requirements)}

\subsection{Task A --- Inheritance + overriding (Exp-09)}
Create \texttt{InheritanceDemo.java}:
\begin{enumerate}[leftmargin=*]
  \item Create base class \texttt{Employee} with fields (\texttt{id}, \texttt{name}) and method \texttt{double calculatePay()}.
  \item Create derived classes:
  \begin{itemize}[leftmargin=*]
    \item \texttt{SalariedEmployee extends Employee} (monthly salary)
    \item \texttt{ContractEmployee extends Employee} (hourlyRate $\times$ hours)
  \end{itemize}
  \item Override \texttt{calculatePay()} in derived classes and use \texttt{@Override}.
  \item Demonstrate runtime polymorphism:
  \begin{itemize}[leftmargin=*]
    \item store mixed objects in an \texttt{Employee[]} array,
    \item call \texttt{calculatePay()} in a loop and print results.
  \end{itemize}
\end{enumerate}

\subsection{Task B --- Abstract class + interface (Exp-10)}
Create \texttt{AbstractInterfaceDemo.java}:
\begin{enumerate}[leftmargin=*]
  \item \textbf{Abstract class part}
  \begin{itemize}[leftmargin=*]
    \item Create abstract class \texttt{EmployeeBase} with an abstract method \texttt{double calculateSalary()}.
    \item Create at least two subclasses (example: \texttt{FullTimeEmployee}, \texttt{InternEmployee}) and implement salary logic.
  \end{itemize}

  \item \textbf{Interface part}
  \begin{itemize}[leftmargin=*]
    \item Create interface \texttt{Payable} with method \texttt{double amountDue()}.
    \item Implement it in two classes (example: \texttt{Invoice} and \texttt{SalarySlip}).
    \item Store objects in a \texttt{Payable[]} array and compute total due.
  \end{itemize}
\end{enumerate}

\section{Hints (read only if stuck)}
\begin{itemize}[leftmargin=*]
  \item Overriding happens when method name + parameters match and the subclass provides a new implementation.
  \item Use \texttt{@Override} to avoid accidental signature mismatches.
  \item Prefer interfaces to model capabilities; use abstract classes when you need shared state + partial implementation.
\end{itemize}

\section{What to submit (MANDATORY)}
Submit a single zipped folder containing:
\begin{itemize}[leftmargin=*]
  \item \texttt{InheritanceDemo.java}
  \item \texttt{AbstractInterfaceDemo.java}
  \item sample output for both programs
  \item observations (5--10 lines): dynamic dispatch; when to prefer interface vs abstract class
\end{itemize}

\notesources{\OOPSUnitIISources}
\end{document}
