\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming (CSEG1044)\\Unit 01 -- Lecture 02 Notes\\Java Overview + Program Structure + Features}
\begin{document}
\maketitle
\tableofcontents

\section{Lecture Snapshot}
\begin{itemize}
  \item \textbf{Unit focus:} Introduction to OOPs and Java Fundamentals
  \item \textbf{Main new ideas:} Java ecosystem: JDK vs JRE vs JVM, Bytecode and the compile/run flow, Basic program structure: class + \texttt{main}
  \item \textbf{Course thread:} Use Java Overview + Program Structure + Features to make the first text-based version of Campus Resource Manager easier to reason about before the full object model arrives.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Lecture Overview}
This lecture makes you comfortable with the Java ``toolchain'' and the minimum you need to compile and run Java programs.
By the end, you should clearly understand:
\begin{itemize}
  \item what the \textbf{JDK}, \textbf{JRE}, and \textbf{JVM} mean,
  \item how \textbf{bytecode} enables Java to run on different platforms,
  \item the difference between \textbf{compile-time errors} and \textbf{runtime errors}.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Prerequisite Bridge}
Before reading the new material in depth, do a fast recall of the older ideas listed below.
\begin{itemize}
  \item \textbf{OOP vocabulary}: The previous lecture introduced objects, classes, and modelling; this lecture connects them to the Java platform. Main reference: \texttt{Unit 01 / Lecture 01 slides/notes}.
  \item \textbf{Problem to program flow}: Once the model is clear, the next question is how Java compiles and runs it. Main reference: \texttt{Unit 01 / Lecture 01 slides/notes}.
\end{itemize}
This lecture only revisits those ideas briefly so that the main explanation can stay focused on the new concept sequence.
\notesources{\OOPSUnitISources}

\section{Learning Outcomes}
After this lecture, you should be able to:
\begin{itemize}
  \item Explain JDK vs JRE vs JVM and what bytecode is.
  \item Compile and run a Java program; recognize compile-time vs runtime errors.
  \item List key features of Java and their practical meaning.
\end{itemize}

\section{Suggested Reading Order}
\begin{enumerate}
  \item Read the \textbf{Prerequisite Bridge} first and confirm each recalled idea.
  \item Study the central explanation in this order: \textit{Java ecosystem: JDK vs JRE vs JVM} $\rightarrow$ \textit{Bytecode and the compile/run flow} $\rightarrow$ \textit{Basic program structure: class + \texttt{main}} $\rightarrow$ \textit{Compile-time vs runtime errors} $\rightarrow$ \textit{Features of Java (what they mean in practice)}.
  \item Trace the worked example line by line before checking short answers.
  \item Attempt the practice section without looking at the solution immediately.
  \item Finish with the exit questions and further-study suggestions to confirm retention.
\end{enumerate}
\notesources{\OOPSUnitISources}

\section{Key Terms (Quick)}
\begin{itemize}
  \item \textbf{JDK (Java Development Kit)}: tools for development (compiler \texttt{javac}, docs, etc.).
  \item \textbf{JRE (Java Runtime Environment)}: runtime needed to run Java apps (JVM + libraries).
  \item \textbf{JVM (Java Virtual Machine)}: executes Java bytecode (\texttt{.class}).
  \item \textbf{Bytecode}: platform-neutral instructions produced by \texttt{javac}.
  \item \textbf{Classpath}: where the JVM looks for compiled classes and libraries.
  \item \textbf{Compile-time error}: detected by the compiler before running.
  \item \textbf{Runtime error/exception}: occurs while the program is executing.
\end{itemize}

\section{Detailed Notes}
\subsection{Java ecosystem: JDK vs JRE vs JVM}
Think of Java in layers:
\begin{itemize}
  \item \textbf{JDK}: used by developers to \emph{create} programs. Includes \texttt{javac} (compiler) and other tools.
  \item \textbf{JRE}: used to \emph{run} programs. Includes the JVM and standard libraries.
  \item \textbf{JVM}: the engine that loads classes, verifies bytecode, and executes it.
\end{itemize}

\paragraph{Rule of thumb.} If you want to \emph{write/compile} Java, install the \textbf{JDK}. If you only want to \emph{run} Java (rare for students), the \textbf{JRE} is enough.

\subsection{Bytecode and the compile/run flow}
Java is often described as ``Write Once, Run Anywhere'' because compilation produces \textbf{bytecode}, not machine code.
The same \texttt{.class} file can run on Windows/Linux/macOS \emph{as long as there is a JVM for that OS}.

\textbf{Flow (mental model):}
\begin{enumerate}
  \item Write: \texttt{Hello.java}
  \item Compile: \texttt{javac Hello.java} $\rightarrow$ \texttt{Hello.class}
  \item Run: \texttt{java Hello} (JVM loads \texttt{Hello.class} and starts \texttt{main})
\end{enumerate}

\subsection{Basic program structure: class + \texttt{main}}
A minimal program has:
\begin{itemize}
  \item a \textbf{class} (container for code),
  \item the \textbf{entry point} method: \texttt{public static void main(String[] args)}.
\end{itemize}

\textbf{Important rule:} if you declare a \texttt{public} class, the file name must match it exactly.
Example: \texttt{public class Hello} must be in \texttt{Hello.java}.

\subsection{Compile-time vs runtime errors}
\textbf{Compile-time error} examples:
\begin{itemize}
  \item syntax errors (missing semicolon, wrong braces),
  \item type errors (assigning \texttt{String} to \texttt{int} without conversion),
  \item missing symbols (calling a method that doesn't exist).
\end{itemize}

\textbf{Runtime error/exception} examples:
\begin{itemize}
  \item \texttt{ArithmeticException} (divide by zero),
  \item \texttt{NullPointerException} (calling method on \texttt{null}),
  \item \texttt{ArrayIndexOutOfBoundsException} (invalid index).
\end{itemize}

\subsection{Features of Java (what they mean in practice)}
\begin{itemize}
  \item \textbf{Platform independent}: bytecode runs on any platform with a JVM.
  \item \textbf{Object-oriented (mostly)}: you model problems using classes/objects (but primitives exist for efficiency).
  \item \textbf{Automatic memory management}: garbage collector reclaims unused objects (no manual \texttt{free}).
  \item \textbf{Strong typing}: many errors are caught at compile time (good for large codebases).
  \item \textbf{Rich standard library}: collections, I/O, networking, GUI, database connectivity, etc.
  \item \textbf{Security model}: class loading + bytecode verification help reduce unsafe behavior.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Running Project Connection}
Use Java Overview + Program Structure + Features to make the first text-based version of Campus Resource Manager easier to reason about before the full object model arrives.
\begin{itemize}
  \item Use Campus Resource Manager as the running case study, or map the same idea to your own project domain if you are using another example.
  \item Ask where today's idea should live: model, service, DAO, UI, or team workflow.
  \item Use the lecture example as a smaller version of the same design choice you will make in the capstone.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Common Mistakes and Confusions}
\begin{itemize}
  \item Using JDK, JRE, and JVM as interchangeable terms.
  \item Forgetting that a public class name must match the file name.
  \item Mixing compile-time errors with runtime failures.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Worked Example: Compile \& Run + Observe Errors}
\subsection*{A minimal program}
\begin{lstlisting}
public class Hello {
    public static void main(String[] args) {
        System.out.println("Hello, Java!");
    }
}
\end{lstlisting}

\subsection*{How to run (terminal)}
\begin{itemize}
  \item Compile: \texttt{javac Hello.java}
  \item Run: \texttt{java Hello}
\end{itemize}

\subsection*{Runtime exception example}
\begin{lstlisting}
public class DivideDemo {
    public static void main(String[] args) {
        int a = 10;
        int b = 0;
        System.out.println(a / b); // ArithmeticException at runtime
    }
}
\end{lstlisting}

\subsection*{Compile-time error example (explain in class)}
If you write \texttt{System.out.println("Hi")} without a semicolon, \texttt{javac} will stop and show an error.
This is a compile-time error: the program never runs.

\section{Demo / Observation Task}
Use this block as a short reproducible demo or observation task.
\begin{itemize}
  \item Reproduce the lecture's main example around \textit{Java ecosystem: JDK vs JRE vs JVM} once without looking at the solution first.
  \item Change one input, rule, or design choice in Campus Resource Manager and predict the result before checking it.
  \item Record one success case, one edge case, and one failure case so the idea becomes observable instead of purely theoretical.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Practice Check (with brief solutions)}
\begin{enumerate}
  \item \textbf{JDK/JRE/JVM:} Which one contains \texttt{javac}? \textbf{Answer:} JDK.
  \item \textbf{File naming:} What should the file name be if you write \texttt{public class Student}? \textbf{Answer:} \texttt{Student.java}.
  \item \textbf{Error type:} Is ``cannot find symbol'' a compile-time or runtime error? \textbf{Answer:} compile-time.
\end{enumerate}

\section{Self-Study Practice (no solutions)}
\begin{enumerate}
  \item Write \texttt{ArgsDemo} that prints all command line arguments with index numbers.
  \item Create 2 small programs: one that fails at compile time, one that fails at runtime. Write a 2-line explanation for each.
  \item Install the JDK and run \texttt{java -version} and \texttt{javac -version}. Note the output.
\end{enumerate}

\section{Further Study}
\begin{itemize}
  \item Revisit the prerequisite references if any recall bullet still feels weak.
  \item Rework the lecture example with one changed assumption, input, or design choice.
  \item Read the textbook and documentation references listed in the source line for a second explanation of the same topic.
  \item Apply the same idea once inside Campus Resource Manager so that the concept moves from theory to design practice.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Exit Questions (with sample answers)}
\textbf{Q1.} Why does Java use bytecode instead of compiling directly to machine code?\par
\textbf{Sample answer:} Bytecode makes the compiled program portable. The same \texttt{.class} can run on different operating systems because the JVM translates/executes bytecode for the current platform.

\vspace{0.6em}
\textbf{Q2.} What is the difference between a compile-time error and a runtime exception?\par
\textbf{Sample answer:} Compile-time errors are detected by \texttt{javac} before execution; runtime exceptions occur while the program is running after compilation succeeds.

\vspace{0.6em}
\textbf{Q3.} Why can one .class file run on multiple operating systems?\par
\textbf{Sample answer:} Because Java compiles to platform-neutral bytecode and each platform provides its own JVM to execute it.

\end{document}
