\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming (CSEG1044)\\Unit 01 -- Lecture 03 Notes\\I/O + Command Line + Types + Operators}
\begin{document}
\maketitle
\tableofcontents

\section{Lecture Snapshot}
\begin{itemize}
  \item \textbf{Unit focus:} Introduction to OOPs and Java Fundamentals
  \item \textbf{Main new ideas:} Console I/O: \texttt{Scanner} (most common for beginners), Console I/O: \texttt{BufferedReader} (fast, line-oriented), Output formatting
  \item \textbf{Course thread:} Use I/O + Command Line + Types + Operators to make the first text-based version of Campus Resource Manager easier to reason about before the full object model arrives.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Lecture Overview}
This lecture focuses on the day-to-day basics you need to write useful Java programs:
\begin{itemize}
  \item taking input (keyboard / terminal),
  \item printing output in a readable way,
  \item using command line arguments,
  \item understanding primitive data types and operators (and common pitfalls).
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Prerequisite Bridge}
Before reading the new material in depth, do a fast recall of the older ideas listed below.
\begin{itemize}
  \item \textbf{Program structure}: You already know where execution starts in Java; this lecture fills that structure with input and expressions. Main reference: \texttt{Unit 01 / Lecture 02 slides/notes}.
  \item \textbf{Basic data handling}: The C prerequisite gives the intuition for values and operators; Java adds stronger typing rules. Main reference: Programming in C prerequisite.
\end{itemize}
This lecture only revisits those ideas briefly so that the main explanation can stay focused on the new concept sequence.
\notesources{\OOPSUnitISources}

\section{Learning Outcomes}
After this lecture, you should be able to:
\begin{itemize}
  \item Read input using \texttt{Scanner} and print formatted output.
  \item Use command line arguments (\texttt{args}) for simple programs.
  \item Apply primitive types, operators, precedence, and casting correctly.
\end{itemize}

\section{Suggested Reading Order}
\begin{enumerate}
  \item Read the \textbf{Prerequisite Bridge} first and confirm each recalled idea.
  \item Study the central explanation in this order: \textit{Console I/O: \texttt{Scanner} (most common for beginners)} $\rightarrow$ \textit{Console I/O: \texttt{BufferedReader} (fast, line-oriented)} $\rightarrow$ \textit{Output formatting} $\rightarrow$ \textit{Command line arguments} $\rightarrow$ \textit{Data types, operators, precedence, casting}.
  \item Trace the worked example line by line before checking short answers.
  \item Attempt the practice section without looking at the solution immediately.
  \item Finish with the exit questions and further-study suggestions to confirm retention.
\end{enumerate}
\notesources{\OOPSUnitISources}

\section{Key Terms (Quick)}
\begin{itemize}
  \item \textbf{Token vs line}: token = one word/number; line = full line of text.
  \item \textbf{Formatting}: printing numbers/text with a chosen format (decimal places, alignment).
  \item \textbf{Primitive types}: built-in value types like \texttt{int}, \texttt{double}, \texttt{char}, \texttt{boolean}.
  \item \textbf{Type casting}: converting a value from one type to another (widening/narrowing).
  \item \textbf{Precedence}: order in which operators are evaluated.
\end{itemize}

\section{Detailed Notes}
\subsection{Console I/O: \texttt{Scanner} (most common for beginners)}
\texttt{Scanner} reads tokens from an input source like \texttt{System.in}.
\begin{itemize}
  \item Use \texttt{nextInt()}, \texttt{nextDouble()}, \texttt{next()} (word), \texttt{nextLine()} (whole line).
  \item \textbf{Common pitfall:} mixing \texttt{nextInt()} and \texttt{nextLine()}.\par
  After reading a number, the newline remains in the buffer; the next \texttt{nextLine()} may return an empty string.\par
  Fix: call an extra \texttt{nextLine()} after reading the number.
\end{itemize}

\subsection{Console I/O: \texttt{BufferedReader} (fast, line-oriented)}
\texttt{BufferedReader} reads full lines as \texttt{String}, and you parse manually:
\begin{itemize}
  \item Good when you want complete control or faster I/O.
  \item Slightly more code than \texttt{Scanner} (because you must parse numbers using \texttt{Integer.parseInt}, etc.).
\end{itemize}

\subsection{Output formatting}
Use:
\begin{itemize}
  \item \texttt{System.out.println(...)} for simple output.
  \item \texttt{System.out.printf("x = \%d, y = \%.2f\%n", x, y);} for formatted output.
\end{itemize}
Common format specifiers:
\begin{itemize}
  \item \texttt{\%d} integer,\;\; \texttt{\%f} floating-point,\;\; \texttt{\%s} string,\;\; \texttt{\%n} newline.
\end{itemize}

\subsection{Command line arguments}
When you run a Java program, anything after the class name becomes a string in \texttt{args}.
\begin{itemize}
  \item Example: \texttt{java Sum 10 20 30}
  \item Then: \texttt{args[0] = "10"}, \texttt{args[1] = "20"}, etc.
\end{itemize}
You often parse numbers using:
\begin{itemize}
  \item \texttt{Integer.parseInt(args[i])}
  \item \texttt{Double.parseDouble(args[i])}
\end{itemize}
\textbf{Pitfall:} if input is not numeric, parsing throws \texttt{NumberFormatException}.

\subsection{Data types, operators, precedence, casting}
\subsubsection{Primitive data types (quick)}
\begin{itemize}
  \item Integer types: \texttt{byte}, \texttt{short}, \texttt{int}, \texttt{long}
  \item Floating types: \texttt{float}, \texttt{double}
  \item Other: \texttt{char} (Unicode), \texttt{boolean}
\end{itemize}

\subsubsection{Operator pitfalls you must know}
\begin{itemize}
  \item \textbf{Integer division:} \texttt{5/2 == 2} (not 2.5). Use \texttt{5/2.0} or cast: \texttt{(double)5/2}.
  \item \textbf{Overflow:} \texttt{int} can overflow silently.\par
  Example: start with \texttt{int x = 2\_000\_000\_000}; then add another billion and observe the wrap-around result.
  \item \textbf{Short-circuit logic:} \texttt{\&\&} and \texttt{||} may skip evaluating the right side.\par
  This is useful for null checks. First test \texttt{s != null}. Only then check whether the string length is greater than zero.
  \item \textbf{Precedence:} multiplication happens before addition. Use parentheses when teaching or learning.
\end{itemize}

\subsubsection{Casting}
\textbf{Widening} (safe) happens automatically: \texttt{int} $\rightarrow$ \texttt{double}.
\textbf{Narrowing} may lose data and requires explicit cast: \texttt{double} $\rightarrow$ \texttt{int}.

\section{Running Project Connection}
Use I/O + Command Line + Types + Operators to make the first text-based version of Campus Resource Manager easier to reason about before the full object model arrives.
\begin{itemize}
  \item Use Campus Resource Manager as the running case study, or map the same idea to your own project domain if you are using another example.
  \item Ask where today's idea should live: model, service, DAO, UI, or team workflow.
  \item Use the lecture example as a smaller version of the same design choice you will make in the capstone.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Common Mistakes and Confusions}
\begin{itemize}
  \item Mixing numeric and line-based input without understanding Scanner behavior.
  \item Ignoring operator precedence and misreading the result.
  \item Expecting Java to convert unrelated types automatically.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Worked Example: Args + Scanner + Formatting}
\begin{lstlisting}
import java.util.Scanner;

public class InputDemo {
    public static void main(String[] args) {
        int maxMarks = 100;
        if (args.length > 0) maxMarks = Integer.parseInt(args[0]);

        Scanner sc = new Scanner(System.in);
        System.out.print("Enter your name: ");
        String name = sc.nextLine();

        System.out.print("Enter marks (0-" + maxMarks + "): ");
        int marks = sc.nextInt();

        double percent = marks * 100.0 / maxMarks;
        System.out.printf("Hello %s, marks=%d/%d, percent=%.2f%%%n",
                name, marks, maxMarks, percent);
        sc.close();
    }
}
\end{lstlisting}

\textbf{What to observe:}
\begin{itemize}
  \item How \texttt{args} is always a string array and needs parsing.
  \item How we used a default value (100) and optionally override it using \texttt{args[0]}.
  \item How formatted output (\texttt{printf}) improves readability.
  \item Why we used \texttt{100.0} in the formula (to avoid integer division mistakes).
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Demo / Observation Task}
Use this block as a short reproducible demo or observation task.
\begin{itemize}
  \item Reproduce the lecture's main example around \textit{Console I/O: \texttt{Scanner} (most common for beginners)} once without looking at the solution first.
  \item Change one input, rule, or design choice in Campus Resource Manager and predict the result before checking it.
  \item Record one success case, one edge case, and one failure case so the idea becomes observable instead of purely theoretical.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Practice Check (with brief solutions)}
\begin{enumerate}
  \item What is the output of \texttt{System.out.println(5/2);}?\; \textbf{Answer:} \texttt{2}.
  \item How to get \texttt{2.5} from 5 and 2?\; \textbf{Answer:} \texttt{5/2.0} or \texttt{(double)5/2}.
  \item If you run \texttt{java Demo A B}, what is \texttt{args.length}?\; \textbf{Answer:} 2.
\end{enumerate}

\section{Self-Study Practice (no solutions)}
\begin{enumerate}
  \item Write a program \texttt{SumArgs} that sums integer command line arguments.
  \item Write a program that reads a line and prints: (a) length, (b) first character, (c) uppercase version.
  \item Create 5 expressions and predict results using precedence; then verify by running Java.
\end{enumerate}

\section{Further Study}
\begin{itemize}
  \item Revisit the prerequisite references if any recall bullet still feels weak.
  \item Rework the lecture example with one changed assumption, input, or design choice.
  \item Read the textbook and documentation references listed in the source line for a second explanation of the same topic.
  \item Apply the same idea once inside Campus Resource Manager so that the concept moves from theory to design practice.
\end{itemize}
\notesources{\OOPSUnitISources}

\section{Exit Questions (with sample answers)}
\textbf{Q1.} Why is \texttt{Scanner.nextLine()} tricky after \texttt{nextInt()}?\par
\textbf{Sample answer:} \texttt{nextInt()} leaves the newline in the input buffer, so the next \texttt{nextLine()} may read the leftover newline and return an empty string. We can consume the newline by calling \texttt{nextLine()} once after reading the number.

\vspace{0.6em}
\textbf{Q2.} Why is \texttt{5/2} not 2.5 in Java?\par
\textbf{Sample answer:} Both operands are integers, so Java performs integer division and discards the fractional part. If either operand is \texttt{double}, it performs floating-point division.

\vspace{0.6em}
\textbf{Q3.} When should you cast explicitly in Java?\par
\textbf{Sample answer:} When Java cannot prove the conversion is safe or when narrowing may lose information.

\end{document}
