\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming Lab (CSEG1044)\\Experiment-07\&08: Classes/Objects + Overloading (Student Handout --- Before Submission)}
\author{Tofik Ali}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

\section{About this handout}
This handout gives the implementation requirements and submission checklist.
Do not copy solutions from outside sources; write and test your own code.

\section{Experiment 07 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To understand classes, objects, and constructors.
  \item \textbf{Problem Statement:} Design classes with data members and member functions.
  \item \textbf{Expected Outcome:} Students will understand object creation and encapsulation.
\end{itemize}

\section{Experiment 08 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To demonstrate compile-time polymorphism.
  \item \textbf{Problem Statement:} Implement method and constructor overloading in Java.
  \item \textbf{Expected Outcome:} Students will understand method resolution at compile time.
\end{itemize}

\section{What you must do (minimum requirements)}

\subsection{Task A --- Student class with validation (Exp-07)}
Create \texttt{StudentDemo.java} with a \texttt{Student} class.

\noindent Fields (private):
\begin{itemize}[leftmargin=*]
  \item \texttt{rollNo} (must be positive)
  \item \texttt{name} (must be non-empty)
  \item \texttt{cgpa} (0 to 10)
\end{itemize}

\noindent Requirements:
\begin{itemize}[leftmargin=*]
  \item Validate in constructor.
  \item Provide a safe update method \texttt{updateCgpa(double)} with validation.
  \item Override \texttt{toString()} to print object details.
\end{itemize}

\subsection{Task B --- Overloading demo (Exp-08)}
Create \texttt{OverloadingDemo.java}:
\begin{enumerate}[leftmargin=*]
  \item Create class \texttt{Box} with overloaded constructors:
  \begin{itemize}[leftmargin=*]
    \item default: unit box
    \item 1 parameter: cube
    \item 3 parameters: \texttt{w, h, d}
  \end{itemize}
  \item Add method \texttt{volume()}.
  \item Write overloaded methods:
  \begin{itemize}[leftmargin=*]
    \item \texttt{area(double r)} $\rightarrow$ circle area
    \item \texttt{area(double w, double h)} $\rightarrow$ rectangle area
  \end{itemize}
  \item In \texttt{main}, create objects using different constructors and print volumes and areas.
\end{enumerate}

\section{Hints (read only if stuck)}
\begin{itemize}[leftmargin=*]
  \item Keep fields private and expose behavior through methods.
  \item Constructor overloading can reduce duplicate initialization code using \texttt{this(...)} chaining.
  \item Overloading cannot be done using only return type; parameter list must change.
\end{itemize}

\section{What to submit (MANDATORY)}
Submit a single zipped folder containing:
\begin{itemize}[leftmargin=*]
  \item \texttt{StudentDemo.java}
  \item \texttt{OverloadingDemo.java}
  \item sample output for both programs
  \item observations (5--10 lines): why encapsulation helps; how Java chooses overloaded methods
\end{itemize}

\notesources{\OOPSUnitIISources}
\end{document}
