\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming (CSEG1044)\\Unit 04 -- Lecture 02 Notes\\Generic Methods + Bounds + Wildcards}
\begin{document}
\maketitle
\tableofcontents

\section{Lecture Snapshot}
\begin{itemize}
  \item \textbf{Unit focus:} Generics, Lambdas, Swing, and JDBC
  \item \textbf{Main new ideas:} Generic methods, Bounds (restrict allowed types), Wildcards: making collection parameters flexible
  \item \textbf{Course thread:} Use Generic Methods + Bounds + Wildcards to move Campus Resource Manager toward a real desktop application with UI, callbacks, and persistence.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Lecture Overview}
In the previous lecture you learned generic classes like \texttt{Box<T>}.
Now we learn three ideas that appear frequently in real Java code:
\begin{itemize}
  \item \textbf{generic methods}: \texttt{<T>} methods that work for many types,
  \item \textbf{bounds}: restricting which types are allowed (\texttt{extends}),
  \item \textbf{wildcards}: flexible method parameters for collections (\texttt{? extends}, \texttt{? super}).
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Prerequisite Bridge}
Before reading the new material in depth, do a fast recall of the older ideas listed below.
\begin{itemize}
  \item \textbf{Generic classes}: This lecture extends the first generics lecture from type declarations to more flexible operations. Main reference: \texttt{Unit 04 / Lecture 01 slides/notes}.
  \item \textbf{Inheritance}: Bounds and wildcards only make sense if subtype relationships are already familiar. Main reference: \texttt{Unit 02 / Lecture 04 slides/notes}.
\end{itemize}
This lecture only revisits those ideas briefly so that the main explanation can stay focused on the new concept sequence.
\notesources{\OOPSUnitIVSources}

\section{Learning Outcomes}
After this lecture, you should be able to:
\begin{itemize}
  \item Write a generic method and explain why it reduces duplication.
  \item Use bounds (\texttt{extends}) to restrict allowed types.
  \item Use wildcards (\texttt{? extends}, \texttt{? super}) with collections.
\end{itemize}

\section{Suggested Reading Order}
\begin{enumerate}
  \item Read the \textbf{Prerequisite Bridge} first and confirm each recalled idea.
  \item Study the central explanation in this order: \textit{Generic methods} $\rightarrow$ \textit{Bounds (restrict allowed types)} $\rightarrow$ \textit{Wildcards: making collection parameters flexible}.
  \item Trace the worked example line by line before checking short answers.
  \item Attempt the practice section without looking at the solution immediately.
  \item Finish with the exit questions and further-study suggestions to confirm retention.
\end{enumerate}
\notesources{\OOPSUnitIVSources}

\section{Key Terms (Quick)}
\begin{itemize}
  \item \textbf{Generic method}: method with its own type parameter, e.g., \texttt{static <T> T first(List<T> list)}.
  \item \textbf{Bound}: restriction on type parameter, e.g., \texttt{<T extends Number>}.
  \item \textbf{Wildcard}: unknown type parameter, e.g., \texttt{List<?>}.
  \item \textbf{Upper-bounded wildcard}: \texttt{? extends Type} (read/produce).
  \item \textbf{Lower-bounded wildcard}: \texttt{? super Type} (write/consume).
  \item \textbf{Invariance}: \texttt{List<Integer>} is not a subtype of \texttt{List<Number>}.
\end{itemize}

\section{Detailed Notes}
\subsection{Generic methods}
Sometimes only one method needs to be generic (not the whole class).
Example: a method that prints any array.

\textbf{Syntax:}
\begin{itemize}
  \item type parameter list comes before return type: \texttt{public static <T> void print(T[] a)}.
\end{itemize}

\subsection{Bounds (restrict allowed types)}
Without bounds, \texttt{T} can be any reference type.
With bounds, you restrict it:
\begin{itemize}
  \item \texttt{<T extends Number>} means \texttt{T} must be \texttt{Number} or a subclass (\texttt{Integer}, \texttt{Double}, ...).
  \item Common pattern: \texttt{<T extends Comparable<T>>} to support comparisons.
\end{itemize}

\subsection{Wildcards: making collection parameters flexible}
The key beginner confusion is:
\begin{itemize}
  \item \texttt{List<Integer>} is \textbf{not} a \texttt{List<Number>} (invariance).
\end{itemize}
So how do we write a method that can accept \texttt{List<Integer>} \emph{or} \texttt{List<Double>}?
We use wildcards.

\subsubsection{\texttt{? extends T} (producer/read)}
If a method only needs to \textbf{read} from a list, use \texttt{? extends}.
Example: sum of numbers:
\begin{itemize}
  \item \texttt{List<? extends Number>} accepts \texttt{List<Integer>}, \texttt{List<Double>}, ...
\end{itemize}
\textbf{Limitation:} you cannot safely add new elements (except \texttt{null}) because the exact subtype is unknown.

\subsubsection{\texttt{? super T} (consumer/write)}
If a method needs to \textbf{add} elements of type \texttt{T} to a list, use \texttt{? super T}.
Example: add integers to a list that can store integers or any of their supertypes.

\paragraph{PECS (memory trick).} Producer Extends, Consumer Super.
\notesources{\OOPSUnitIVSources}

\section{Running Project Connection}
Use Generic Methods + Bounds + Wildcards to move Campus Resource Manager toward a real desktop application with UI, callbacks, and persistence.
\begin{itemize}
  \item Use Campus Resource Manager as the running case study, or map the same idea to your own project domain if you are using another example.
  \item Ask where today's idea should live: model, service, DAO, UI, or team workflow.
  \item Use the lecture example as a smaller version of the same design choice you will make in the capstone.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Common Mistakes and Confusions}
\begin{itemize}
  \item Using a wildcard when a named type parameter would be clearer.
  \item Overengineering beginner APIs with too many bounds.
  \item Assuming List<Child> is automatically a List<Parent>.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Worked Example: generic methods + bounds + wildcards}
\subsection*{Generic method (print array)}
\begin{lstlisting}
public class GenericDemo {
    public static <T> void printArray(T[] a) {
        for (T x : a) System.out.print(x + " ");
        System.out.println();
    }
}
\end{lstlisting}

\subsection*{Bounded type parameter (max)}
\begin{lstlisting}
public static <T extends Comparable<T>> T max(T a, T b) {
    return (a.compareTo(b) >= 0) ? a : b;
}
\end{lstlisting}

\subsection*{Upper-bounded wildcard (sum)}
\begin{lstlisting}
import java.util.List;

public static double sum(List<? extends Number> list) {
    double s = 0.0;
    for (Number n : list) s += n.doubleValue();
    return s;
}
\end{lstlisting}

\subsection*{Lower-bounded wildcard (add)}
\begin{lstlisting}
import java.util.List;

public static void addOnes(List<? super Integer> list, int count) {
    for (int i = 0; i < count; i++) list.add(1);
}
\end{lstlisting}

\section{Demo / Observation Task}
Use this block as a short reproducible demo or observation task.
\begin{itemize}
  \item Reproduce the lecture's main example around \textit{Generic methods} once without looking at the solution first.
  \item Change one input, rule, or design choice in Campus Resource Manager and predict the result before checking it.
  \item Record one success case, one edge case, and one failure case so the idea becomes observable instead of purely theoretical.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Practice Check (with brief solutions)}
\begin{enumerate}
  \item Why is \texttt{List<Integer>} not a subtype of \texttt{List<Number>}?\par
  \textbf{Answer:} generics are invariant in Java; allowing it would break type safety (you could add a \texttt{Double} into a \texttt{List<Integer>} through a \texttt{List<Number>} reference).

  \item If you only need to read from a list, which wildcard is common: \texttt{extends} or \texttt{super}?\par
  \textbf{Answer:} \texttt{? extends ...}.

  \item What is the PECS rule?\par
  \textbf{Answer:} Producer Extends, Consumer Super.
\end{enumerate}

\section{Self-Study Practice (no solutions)}
\begin{enumerate}
  \item Write a generic method \texttt{swap(T[] a, int i, int j)}.
  \item Write \texttt{min} using \texttt{<T extends Comparable<T>>}.
  \item Write a method that copies from one list to another using wildcards (hint: one produces, one consumes).
\end{enumerate}

\section{Further Study}
\begin{itemize}
  \item Revisit the prerequisite references if any recall bullet still feels weak.
  \item Rework the lecture example with one changed assumption, input, or design choice.
  \item Read the textbook and documentation references listed in the source line for a second explanation of the same topic.
  \item Apply the same idea once inside Campus Resource Manager so that the concept moves from theory to design practice.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Exit Questions (with sample answers)}
\textbf{Q1.} When do you use \texttt{? extends T}?\par
\textbf{Sample answer:} Use \texttt{? extends T} when your method only needs to read values of type \texttt{T} from a structure and doesn't need to add elements, because it can accept multiple subtype lists safely.

\vspace{0.6em}
\textbf{Q2.} What does a bound like \texttt{<T extends Number>} mean?\par
\textbf{Sample answer:} It means the type parameter \texttt{T} must be \texttt{Number} or a subclass of \texttt{Number}, so you can safely use \texttt{Number}-level operations inside the method/class.

\vspace{0.6em}
\textbf{Q3.} What problem do bounds solve in generic APIs?\par
\textbf{Sample answer:} They restrict allowed types so the compiler can guarantee valid operations inside the API.

\end{document}
