\documentclass{beamer}

\input{../../../_shared/beamer-common.tex}

\title[OOPS with Java]{Object Oriented Programming (CSEG1044)}
\subtitle{Unit 01 -- Lecture 04: Control Flow + Arrays + Strings}
\begin{document}

\begin{frame}[fragile]
  \titlepage
  \vspace{0.5em}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Quick Links}
  \centering
  \hyperlink{sec:overview}{\beamerbutton{Overview}}\hspace{0.6em}
  \hyperlink{sec:concepts}{\beamerbutton{Key Topics}}\hspace{0.6em}
  \hyperlink{sec:demo}{\beamerbutton{Demo}}\hspace{0.6em}
  \hyperlink{sec:summary}{\beamerbutton{Summary}}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Agenda}
  \tableofcontents
  \slidesources{\OOPSUnitISources}
\end{frame}

\section{Overview}
\label{sec:overview}

\begin{frame}{Lecture Snapshot}
\begin{itemize}[<+->]
  \item \textbf{Unit focus:} Introduction to OOPs and Java Fundamentals
  \item \textbf{Course thread:} Use Control Flow + Arrays + Strings to make the first text-based version of Campus Resource Manager easier to reason about before the full object model arrives.
\end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Recall Before You Start}
\begin{itemize}[<+->]
  \item \textbf{Types and operators}: Conditions, loop bounds, indices, and string logic depend on the previous lecture. Main reference: \texttt{Unit 01 / Lecture 03 slides/notes}.
  \item \textbf{Program flow}: Control statements only make sense once you already know the default top-to-bottom execution order. Main reference: \texttt{Unit 01 / Lecture 02 slides/notes}.
\end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}


\begin{frame}{Learning Outcomes}
  \begin{itemize}[<+->]
    \item Use \texttt{if/else}, \texttt{switch}, and loops to control program flow.
    \item Create/traverse 1D and 2D arrays and solve simple array problems.
    \item Explain \texttt{String} immutability and use \texttt{StringBuilder} when needed.
  \end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Key Terms to Know}
\small
\begin{columns}[t]
\column{0.48\textwidth}
\begin{itemize}
  \item \textbf{Selection}: choosing one path among many (\texttt{if}, \texttt{switch}).
  \item \textbf{Iteration}: repeating work until a condition changes (\texttt{for}, \texttt{while}).
  \item \textbf{Array}: fixed-size contiguous collection of same-type elements.
\end{itemize}
\column{0.48\textwidth}
\begin{itemize}
  \item \textbf{Index}: position in array (0-based in Java).
  \item \textbf{Immutable}: cannot be changed after creation (e.g., \texttt{String}).
  \item \textbf{StringBuilder}: mutable sequence of characters (efficient for repeated concatenation).
\end{itemize}
\end{columns}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Study Flow for This Lecture}
\begin{enumerate}[<+->]
  \item Refresh the prerequisite bridge before reading new ideas in isolation.
  \item Study the main build in this order: \textit{Program control statements (selection + loops)} $\rightarrow$ \textit{Arrays (1D/2D), types of arrays, common patterns} $\rightarrow$ \textit{Strings and \texttt{StringBuilder}}.
  \item Trace the worked example and connect each line back to one concept frame.
  \item Attempt the mini exercises before moving to the recap and exit check.
\end{enumerate}
  \slidesources{\OOPSUnitISources}
\end{frame}


\section{Key Topics}
\label{sec:concepts}

\begin{frame}{Unit 01: Introduction to OOPs}
  \begin{itemize}[<+->]
    \item Lecture focus: Control Flow + Arrays + Strings
  \end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Today: Roadmap}
  \begin{itemize}[<+->]
    \item Program control statements (selection + loops)
    \item Arrays (1D/2D), types of arrays, common patterns
    \item Strings (immutability, common methods), `StringBuilder` (need-based)
  \end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Control flow (quick)}
  \begin{itemize}[<+->]
    \item Selection: \texttt{if/else}, \texttt{switch}
    \item Loops: \texttt{for}, \texttt{while}, \texttt{do-while}
    \item Useful keywords: \texttt{break}, \texttt{continue}
  \end{itemize}
  \begin{block}{Common pitfalls}
    \begin{itemize}
      \item Off-by-one errors in loops
      \item Missing \texttt{break} in \texttt{switch} (unintended fall-through)
    \end{itemize}
  \end{block}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Arrays (1D/2D)}
  \begin{itemize}[<+->]
    \item 1D: \texttt{int[] a = new int[n];} and \texttt{a.length}
    \item Traversal: \texttt{for (int i=0; i<a.length; i++)} or \texttt{for (int x: a)}
    \item 2D: \texttt{int[][] m = new int[r][c];}
    \item Jagged arrays: each row can have different length (use \texttt{m[i].length})
  \end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Strings (must-know)}
  \begin{itemize}[<+->]
    \item \texttt{String} is \textbf{immutable}: methods return new strings.
    \item Compare content with \texttt{equals()}, not \texttt{==}.
    \item Repeated concatenation in loops is slow $\rightarrow$ use \texttt{StringBuilder}.
  \end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Checkpoint and Reflection}
\begin{block}{Reflection prompt}
Where would \textit{Control Flow + Arrays + Strings} matter most in Campus Resource Manager, and which design choice would you justify using this lecture's ideas?
\end{block}
\begin{block}{Quick self-check}
Which part needs one more example before you move on: \textit{Program control statements (selection + loops)}, \textit{Arrays (1D/2D), types of arrays, common patterns}, the worked example, or the recap?
\end{block}
  \slidesources{\OOPSUnitISources}
\end{frame}

\section{Demo / Example}
\label{sec:demo}

\begin{frame}[fragile,shrink=8]{Example: grades (arrays + control flow) (1/2)}
\begin{lstlisting}
import java.util.Scanner;

public class GradeStats {
    public static void main(String[] args) {
        Scanner sc = new Scanner(System.in);
        System.out.print("Enter number of students: ");
        int n = sc.nextInt();

        int[] marks = new int[n];
        for (int i = 0; i < n; i++) {
            System.out.print("Marks for student " + (i + 1) + ": ");
            marks[i] = sc.nextInt();
        }
\end{lstlisting}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}[fragile]{Example: grades (arrays + control flow) (2/2)}
\begin{lstlisting}
// continued...
        int max = marks[0], sum = 0;
        for (int x : marks) { sum += x; if (x > max) max = x; }
        double avg = sum / (double) n;
        System.out.printf("Max=%d, Avg=%.2f%n", max, avg);
        sc.close();
    }
}
\end{lstlisting}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Mini Exercises (2--3 mins)}
  \begin{enumerate}[<+->]
    \item Array of length 10: what is the last valid index?
    \item Which compares string content: \texttt{==} or \texttt{equals}?
    \item Which loop runs at least once: \texttt{while} or \texttt{do-while}?
  \end{enumerate}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Watch-outs}
\begin{itemize}[<+->]
  \item Writing off-by-one loop boundaries.
  \item Comparing String values with == instead of an equality method.
  \item Treating String as mutable and expecting in-place updates.
\end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}

\section{Summary}
\label{sec:summary}

\begin{frame}{Key Takeaways}
  \begin{itemize}[<+->]
    \item Use selection + loops to express logic clearly (avoid off-by-one bugs).
    \item Arrays store many values of the same type; \texttt{length} guides safe iteration.
    \item Strings are immutable; use \texttt{equals()} and \texttt{StringBuilder} when needed.
  \end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Checkpoint Questions}
  \begin{enumerate}[<+->]
    \item How do you avoid off-by-one errors in array loops?
    \item Why should we use \texttt{equals()} for strings?
    \item When is \texttt{StringBuilder} preferred over \texttt{String} concatenation?
  \end{enumerate}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Next Study Steps}
\begin{itemize}[<+->]
  \item Revisit the recall references if any older idea still feels weak.
  \item Rework the lecture example without looking at the notes first.
  \item Attempt the self-study practice and then answer the exit questions in full sentences.
  \item Apply the idea once inside Campus Resource Manager to make the concept stick.
\end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}

\begin{frame}{Next Unit Preview}
  \textbf{Next:} Unit 02 -- Classes, Objects, and Encapsulation
  \vspace{0.6em}
  \begin{itemize}[<+->]
    \item We will start writing proper classes with private fields and methods.
    \item Bring one example of a real-world entity you want to model as a class.
  \end{itemize}
  \slidesources{\OOPSUnitISources}
\end{frame}

\end{document}
