\documentclass{beamer}

\input{../../../_shared/beamer-common.tex}

\title[OOPS with Java]{Object Oriented Programming (CSEG1044)}
\subtitle{Unit 02 -- Lecture 01: Classes, Objects, and Encapsulation (Basics)}
\begin{document}

\begin{frame}[fragile]
  \titlepage
  \vspace{0.5em}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Quick Links}
  \centering
  \hyperlink{sec:overview}{\beamerbutton{Overview}}\hspace{0.6em}
  \hyperlink{sec:concepts}{\beamerbutton{Key Topics}}\hspace{0.6em}
  \hyperlink{sec:demo}{\beamerbutton{Demo}}\hspace{0.6em}
  \hyperlink{sec:summary}{\beamerbutton{Summary}}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Agenda}
  \tableofcontents
  \slidesources{\OOPSUnitIISources}
\end{frame}

\section{Overview}
\label{sec:overview}

\begin{frame}{Lecture Snapshot}
\begin{itemize}[<+->]
  \item \textbf{Unit focus:} Classes, Inheritance, Packages, and Interfaces
  \item \textbf{Course thread:} Use Classes, Objects, and Encapsulation (Basics) to strengthen the domain model, APIs, and access rules inside Campus Resource Manager.
\end{itemize}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Recall Before You Start}
\begin{itemize}[<+->]
  \item \textbf{Encapsulation principle}: Unit 01 introduced encapsulation conceptually; this lecture turns it into actual Java class design. Main reference: \texttt{Unit 01 / Lecture 01 slides/notes}.
  \item \textbf{Java class syntax}: You already saw the minimum class structure and now reuse it for reusable domain models. Main reference: \texttt{Unit 01 / Lecture 02 slides/notes}.
\end{itemize}
  \slidesources{\OOPSUnitIISources}
\end{frame}


\begin{frame}{Learning Outcomes}
  \begin{itemize}[<+->]
    \item Identify fields vs methods and explain state vs behavior.
    \item Explain references and aliasing using a small example.
    \item Apply encapsulation using \texttt{private} fields and validation methods.
  \end{itemize}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Key Terms to Know}
\small
\begin{columns}[t]
\column{0.48\textwidth}
\begin{itemize}
  \item \textbf{Field}: data stored inside an object (state).
  \item \textbf{Method}: behavior (operations) of an object.
  \item \textbf{Reference}: a variable that points to an object (not the object itself).
\end{itemize}
\column{0.48\textwidth}
\begin{itemize}
  \item \textbf{Aliasing}: two references pointing to the same object.
  \item \textbf{Encapsulation}: keeping fields private and changing state through methods.
  \item \textbf{Access modifier}: \texttt{private}, default, \texttt{protected}, \texttt{public} (controls visibility).
\end{itemize}
\end{columns}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Study Flow for This Lecture}
\begin{enumerate}[<+->]
  \item Refresh the prerequisite bridge before reading new ideas in isolation.
  \item Study the main build in this order: \textit{Class fundamentals: fields, methods, state vs behavior} $\rightarrow$ \textit{Objects and reference semantics (mental model)} $\rightarrow$ \textit{Encapsulation: protect invariants}.
  \item Trace the worked example and connect each line back to one concept frame.
  \item Attempt the mini exercises before moving to the recap and exit check.
\end{enumerate}
  \slidesources{\OOPSUnitIISources}
\end{frame}


\section{Key Topics}
\label{sec:concepts}

\begin{frame}{Unit 02: Classes, Inheritance, Packages and Interfaces}
  \begin{itemize}[<+->]
    \item Lecture focus: Classes, Objects, and Encapsulation (Basics)
  \end{itemize}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Today: Roadmap}
  \begin{itemize}[<+->]
    \item Class fundamentals: fields, methods, state vs behavior
    \item Objects and reference semantics; object diagrams (mental model)
    \item Encapsulation motivation (invariants) + intro to access modifiers
  \end{itemize}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Class vs Object (quick)}
  \begin{itemize}[<+->]
    \item \textbf{Class} = blueprint (type): fields + methods
    \item \textbf{Object} = instance at runtime (identity + state + behavior)
    \item State = values in fields; Behavior = what methods do
  \end{itemize}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{References and aliasing (important)}
  \begin{itemize}[<+->]
    \item In Java, variables of class type hold \textbf{references}.
    \item Assignment copies the reference (not the whole object).
    \item If two references point to the same object, changes are shared (\textbf{aliasing}).
  \end{itemize}
  \begin{block}{Why it matters}
    Many bugs come from unexpected shared updates.
  \end{block}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Encapsulation = protect invariants}
  \begin{itemize}[<+->]
    \item Make fields \texttt{private}.
    \item Update state via methods that validate inputs.
    \item Keep object invariants always true (e.g., \texttt{0 <= cgpa <= 10}).
  \end{itemize}
  \begin{block}{Common mistake}
    Adding setters for everything and allowing invalid values.
  \end{block}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Checkpoint and Reflection}
\begin{block}{Reflection prompt}
Where would \textit{Classes, Objects, and Encapsulation (Basics)} matter most in Campus Resource Manager, and which design choice would you justify using this lecture's ideas?
\end{block}
\begin{block}{Quick self-check}
Which part needs one more example before you move on: \textit{Class fundamentals: fields, methods, state vs behavior}, \textit{Objects and reference semantics (mental model)}, the worked example, or the recap?
\end{block}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\section{Demo / Example}
\label{sec:demo}

\begin{frame}[fragile,shrink=8]{Example: \texttt{Student} (encapsulation) (1/2)}
\begin{lstlisting}
public class Student {
    private final String rollNo;
    private final String name;
    private double cgpa; // invariant: 0..10

    public Student(String rollNo, String name, double cgpa) {
        if (cgpa < 0.0 || cgpa > 10.0)
            throw new IllegalArgumentException("cgpa");
        this.rollNo = rollNo;
        this.name = name;
        this.cgpa = cgpa;
    }
\end{lstlisting}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}[fragile]{Example: \texttt{Student} (encapsulation) (2/2)}
\begin{lstlisting}
// continued...

    public void updateCgpa(double newCgpa) {
        if (newCgpa < 0.0 || newCgpa > 10.0)
            throw new IllegalArgumentException("cgpa");
        cgpa = newCgpa;
    }

    public double getCgpa() { return cgpa; }
}
\end{lstlisting}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}[fragile]{Aliasing demo (observe)}
\begin{lstlisting}
public class Demo {
    public static void main(String[] args) {
        Student a = new Student("21CS1001", "Asha", 8.2);
        Student b = a;   // same object
        b.updateCgpa(9.0);
        System.out.println(a.getCgpa()); // prints 9.0
    }
}
\end{lstlisting}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Mini Exercises (2--3 mins)}
  \begin{enumerate}[<+->]
    \item What is the difference between a class and an object?
    \item If \texttt{b = a} and \texttt{b} modifies the object, does \texttt{a} see it?
    \item Give one invariant for \texttt{BankAccount} or \texttt{LibraryBook}.
  \end{enumerate}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Watch-outs}
\begin{itemize}[<+->]
  \item Making every field public for convenience.
  \item Confusing an object reference with a copy of the object state.
  \item Thinking classes matter only inside main.
\end{itemize}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\section{Summary}
\label{sec:summary}

\begin{frame}{Key Takeaways}
  \begin{itemize}[<+->]
    \item Classes define state + behavior; objects are runtime instances.
    \item Java object variables are references; aliasing is common.
    \item Encapsulation protects invariants using \texttt{private} + validation methods.
  \end{itemize}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Checkpoint Questions}
  \begin{enumerate}[<+->]
    \item Why do we keep fields \texttt{private}?
    \item What is aliasing and why can it be risky?
    \item What is an invariant? Give one example.
  \end{enumerate}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Next Study Steps}
\begin{itemize}[<+->]
  \item Revisit the recall references if any older idea still feels weak.
  \item Rework the lecture example without looking at the notes first.
  \item Attempt the self-study practice and then answer the exit questions in full sentences.
  \item Apply the idea once inside Campus Resource Manager to make the concept stick.
\end{itemize}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\begin{frame}{Next Lecture Preview}
  \textbf{Next:} Constructors, \texttt{this}, Overloading, \texttt{static}, and GC overview
  \vspace{0.6em}
  \begin{itemize}[<+->]
    \item We will learn how objects are created and how constructors enforce invariants.
    \item We will also learn when to use \texttt{static}.
  \end{itemize}
  \slidesources{\OOPSUnitIISources}
\end{frame}

\end{document}
