\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming (CSEG1044)\\Unit 02 -- Lecture 05 Notes\\Abstract Classes, `final`, and Interfaces (Design Choices)}
\begin{document}
\maketitle
\tableofcontents

\section{Lecture Snapshot}
\begin{itemize}
  \item \textbf{Unit focus:} Classes, Inheritance, Packages, and Interfaces
  \item \textbf{Main new ideas:} Abstract classes and methods, \texttt{final} keyword (design intent), Interfaces: built-in + user-defined
  \item \textbf{Course thread:} Use Abstract Classes, `final`, and Interfaces (Design Choices) to strengthen the domain model, APIs, and access rules inside Campus Resource Manager.
\end{itemize}
\notesources{\OOPSUnitIISources}

\section{Lecture Overview}
By now you have seen basic classes, inheritance, and an interface intro.
This lecture answers the practical design questions:
\begin{itemize}
  \item When should I use an \textbf{abstract class}?
  \item When should I use an \textbf{interface}?
  \item When should something be \textbf{final}?
\end{itemize}
\notesources{\OOPSUnitIISources}

\section{Prerequisite Bridge}
Before reading the new material in depth, do a fast recall of the older ideas listed below.
\begin{itemize}
  \item \textbf{Inheritance}: Abstract classes extend inheritance by allowing an incomplete but shared parent design. Main reference: \texttt{Unit 02 / Lecture 04 slides/notes}.
  \item \textbf{Interface idea}: You already saw interfaces as contracts; this lecture compares them with abstract classes in more depth. Main reference: \texttt{Unit 02 / Lecture 03 slides/notes}.
\end{itemize}
This lecture only revisits those ideas briefly so that the main explanation can stay focused on the new concept sequence.
\notesources{\OOPSUnitIISources}

\section{Learning Outcomes}
After this lecture, you should be able to:
\begin{itemize}
  \item Use abstract classes for shared code + required methods.
  \item Explain \texttt{final} for variables/methods/classes.
  \item Use interfaces (including \texttt{Comparable}, \texttt{Runnable}) as contracts.
\end{itemize}

\section{Suggested Reading Order}
\begin{enumerate}
  \item Read the \textbf{Prerequisite Bridge} first and confirm each recalled idea.
  \item Study the central explanation in this order: \textit{Abstract classes and methods} $\rightarrow$ \textit{\texttt{final} keyword (design intent)} $\rightarrow$ \textit{Interfaces: built-in + user-defined}.
  \item Trace the worked example line by line before checking short answers.
  \item Attempt the practice section without looking at the solution immediately.
  \item Finish with the exit questions and further-study suggestions to confirm retention.
\end{enumerate}
\notesources{\OOPSUnitIISources}

\section{Key Terms (Quick)}
\begin{itemize}
  \item \textbf{Abstract class}: class that cannot be instantiated; may include abstract methods.
  \item \textbf{Abstract method}: declared without body; must be implemented by concrete subclasses.
  \item \textbf{Interface}: contract specifying methods a class must provide.
  \item \textbf{Final}: ``cannot change'' (variable), ``cannot override'' (method), ``cannot extend'' (class).
  \item \textbf{Contract}: what is promised by a type (what callers can rely on).
\end{itemize}

\section{Detailed Notes}
\subsection{Abstract classes and methods}
Use an abstract class when:
\begin{itemize}
  \item you want to enforce a common API (method signatures),
  \item and you also want to share some implementation code across subclasses.
\end{itemize}

\textbf{Key properties:}
\begin{itemize}
  \item Cannot instantiate: \texttt{new Shape()} is illegal if \texttt{Shape} is abstract.
  \item Subclasses must implement all abstract methods (unless they are also abstract).
  \item An abstract class can contain concrete fields and methods.
\end{itemize}

\subsection{\texttt{final} keyword (design intent)}
\textbf{Final variable:} assigned once.
\begin{itemize}
  \item For primitives: value cannot change.
  \item For references: reference cannot change, but the object may still change unless the object is immutable.
\end{itemize}

\textbf{Final method:} cannot be overridden.
\begin{itemize}
  \item Used to lock down behavior and prevent subclasses from changing critical logic.
\end{itemize}

\textbf{Final class:} cannot be extended.
\begin{itemize}
  \item Example: \texttt{String} is final (one reason is security/immutability design).
\end{itemize}

\subsection{Interfaces: built-in + user-defined}
Use an interface when:
\begin{itemize}
  \item you want multiple unrelated classes to share a capability (e.g., \texttt{Comparable}),
  \item you want multiple implementations (e.g., different payment methods),
  \item you want flexible designs with low coupling.
\end{itemize}

\textbf{Built-in examples:}
\begin{itemize}
  \item \texttt{Comparable<T>}: defines natural ordering (\texttt{compareTo}). Used by sorting.
  \item \texttt{Runnable}: defines a task that can run in a thread (\texttt{run}).
\end{itemize}
\notesources{\OOPSUnitIISources}

\section{Running Project Connection}
Use Abstract Classes, `final`, and Interfaces (Design Choices) to strengthen the domain model, APIs, and access rules inside Campus Resource Manager.
\begin{itemize}
  \item Use Campus Resource Manager as the running case study, or map the same idea to your own project domain if you are using another example.
  \item Ask where today's idea should live: model, service, DAO, UI, or team workflow.
  \item Use the lecture example as a smaller version of the same design choice you will make in the capstone.
\end{itemize}
\notesources{\OOPSUnitIISources}

\section{Common Mistakes and Confusions}
\begin{itemize}
  \item Trying to instantiate an abstract class directly.
  \item Marking too many classes or methods final too early.
  \item Choosing an interface even when shared state or shared code is central.
\end{itemize}
\notesources{\OOPSUnitIISources}

\section{Worked Example: \texttt{Comparable} for sorting}
\begin{lstlisting}
import java.util.Arrays;

class Student implements Comparable<Student> {
    private final String rollNo;
    private final double cgpa;

    Student(String rollNo, double cgpa) {
        this.rollNo = rollNo;
        this.cgpa = cgpa;
    }

    @Override
    public int compareTo(Student other) {
        // descending order by cgpa
        return Double.compare(other.cgpa, this.cgpa);
    }

    @Override
    public String toString() {
        return rollNo + ":" + cgpa;
    }
}

public class Demo {
    public static void main(String[] args) {
        Student[] a = {
            new Student("21CS1001", 8.2),
            new Student("21CS1002", 9.1),
            new Student("21CS1003", 7.6)
        };
        Arrays.sort(a); // uses compareTo
        System.out.println(Arrays.toString(a));
    }
}
\end{lstlisting}

\section{Demo / Observation Task}
Use this block as a short reproducible demo or observation task.
\begin{itemize}
  \item Reproduce the lecture's main example around \textit{Abstract classes and methods} once without looking at the solution first.
  \item Change one input, rule, or design choice in Campus Resource Manager and predict the result before checking it.
  \item Record one success case, one edge case, and one failure case so the idea becomes observable instead of purely theoretical.
\end{itemize}
\notesources{\OOPSUnitIISources}

\section{Practice Check (with brief solutions)}
\begin{enumerate}
  \item Can we create an object of an abstract class?\; \textbf{Answer:} No.
  \item What does \texttt{final class} mean?\; \textbf{Answer:} cannot be extended.
  \item Why do we use interfaces?\; \textbf{Answer:} contracts + low coupling + multiple implementations.
\end{enumerate}

\section{Self-Study Practice (no solutions)}
\begin{enumerate}
  \item Create an abstract class \texttt{Shape} with abstract method \texttt{area()} and implement \texttt{Circle} and \texttt{Rectangle}.
  \item Create an interface \texttt{Payable} and implement it for \texttt{Invoice} and \texttt{SalarySlip}.
  \item Implement \texttt{Runnable} in a class \texttt{HelloTask} and print a message in \texttt{run()} (we will start threads in Unit 03).
\end{enumerate}

\section{Further Study}
\begin{itemize}
  \item Revisit the prerequisite references if any recall bullet still feels weak.
  \item Rework the lecture example with one changed assumption, input, or design choice.
  \item Read the textbook and documentation references listed in the source line for a second explanation of the same topic.
  \item Apply the same idea once inside Campus Resource Manager so that the concept moves from theory to design practice.
\end{itemize}
\notesources{\OOPSUnitIISources}

\section{Exit Questions (with sample answers)}
\textbf{Q1.} When do you prefer an interface over an abstract class?\par
\textbf{Sample answer:} Prefer an interface when you want a contract that many unrelated classes can implement and you want flexibility and low coupling. Use abstract class when you also want to share code and common state.

\vspace{0.6em}
\textbf{Q2.} What is the meaning of \texttt{final} in different contexts?\par
\textbf{Sample answer:} A final variable cannot be reassigned, a final method cannot be overridden, and a final class cannot be extended.

\vspace{0.6em}
\textbf{Q3.} When is an interface better than an abstract class?\par
\textbf{Sample answer:} When you want a loose contract, multiple implementations, and no required shared state.

\end{document}
