\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming Lab (CSEG1044)\\Experiment-13\&14: File I/O + Generics (Student Handout --- Before Submission)}
\author{Tofik Ali}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

\section{About this handout}
This handout specifies the required programs for file handling and generics.
Write and test your own implementation.

\section{Experiment 13 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To perform file operations using Java I/O streams.
  \item \textbf{Problem Statement:} Develop programs using \texttt{FileReader} and \texttt{FileWriter}.
  \item \textbf{Expected Outcome:} Students will handle persistent data.
\end{itemize}

\section{Experiment 14 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To implement generic classes and methods.
  \item \textbf{Problem Statement:} Design programs using generics.
  \item \textbf{Expected Outcome:} Students will understand type safety.
\end{itemize}

\section{What you must do (minimum requirements)}

\subsection{Task A --- File I/O (Exp-13)}
Create \texttt{FileIOStreamsDemo.java}:
\begin{enumerate}[leftmargin=*]
  \item Write 3 lines to \texttt{input.txt} using \texttt{FileWriter}.
  \item Read and print the file contents using \texttt{FileReader}.
  \item Copy \texttt{input.txt} to \texttt{copy.txt} using a character buffer.
  \item Print how many characters were copied.
  \item Use try-with-resources and handle \texttt{IOException}.
\end{enumerate}

\subsection{Task B --- Generics basics (Exp-14)}
Create \texttt{GenericsDemo.java}:
\begin{enumerate}[leftmargin=*]
  \item Create a generic class \texttt{Box<T>} that stores a single value and returns it via \texttt{get()}.
  \item Create a generic method \texttt{printTwice(T x)} and call it with different types.
  \item Demonstrate at least:
  \begin{itemize}[leftmargin=*]
    \item \texttt{Box<Integer>}
    \item \texttt{Box<String>}
  \end{itemize}
  \item Optional: create a generic \texttt{Pair<K,V>} and print it.
\end{enumerate}

\section{Hints (read only if stuck)}
\begin{itemize}[leftmargin=*]
  \item Use try-with-resources to auto-close \texttt{Reader/Writer}.
  \item Relative file paths are resolved from your program's working directory (often your project folder).
  \item Generics prevent many \texttt{ClassCastException} issues by enforcing type checks at compile time.
  \item Avoid raw types (use \texttt{Box<String>} instead of \texttt{Box}).
\end{itemize}

\section{What to submit (MANDATORY)}
Submit a single zipped folder containing:
\begin{itemize}[leftmargin=*]
  \item \texttt{FileIOStreamsDemo.java}
  \item \texttt{GenericsDemo.java}
  \item sample output for both programs
  \item observations (5--10 lines): what try-with-resources does; where generics helped
\end{itemize}

\notesources{\OOPSUnitIIISources\ \OOPSUnitIVSources}
\end{document}
