\documentclass{beamer}

\input{../../../_shared/beamer-common.tex}

\title[OOPS with Java]{Object Oriented Programming (CSEG1044)}
\subtitle{Unit 03 -- Lecture 02: Exception Handling (Core)}
\begin{document}

\begin{frame}[fragile]
  \titlepage
  \vspace{0.5em}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Quick Links}
  \centering
  \hyperlink{sec:overview}{\beamerbutton{Overview}}\hspace{0.6em}
  \hyperlink{sec:concepts}{\beamerbutton{Key Topics}}\hspace{0.6em}
  \hyperlink{sec:demo}{\beamerbutton{Demo}}\hspace{0.6em}
  \hyperlink{sec:summary}{\beamerbutton{Summary}}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Agenda}
  \tableofcontents
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\section{Overview}
\label{sec:overview}

\begin{frame}{Lecture Snapshot}
\begin{itemize}[<+->]
  \item \textbf{Unit focus:} Nested Classes, Exceptions, Multithreading, and I/O
  \item \textbf{Course thread:} Use Exception Handling (Core) to make Campus Resource Manager safer, more modular, and easier to recover when something goes wrong.
\end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Recall Before You Start}
\begin{itemize}[<+->]
  \item \textbf{Control flow}: try, catch, and finally are still flow-control ideas for abnormal situations. Main reference: \texttt{Unit 01 / Lecture 04 slides/notes}.
  \item \textbf{Validation design}: Encapsulated objects should reject invalid state changes in a disciplined way. Main reference: \texttt{Unit 02 / Lecture 01 slides/notes}.
\end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}


\begin{frame}{Learning Outcomes}
  \begin{itemize}[<+->]
    \item Distinguish exception vs error; checked vs unchecked.
    \item Write \texttt{try/catch/finally} and use \texttt{throw} vs \texttt{throws}.
    \item Design a small custom exception for validation.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Key Terms to Know}
\small
\begin{columns}[t]
\column{0.48\textwidth}
\begin{itemize}
  \item \textbf{Exception}: abnormal condition a program can potentially handle.
  \item \textbf{Error}: serious problem typically not recoverable (e.g., \texttt{OutOfMemoryError}).
  \item \textbf{Checked exception}: must be caught or declared (e.g., \texttt{IOException}).
\end{itemize}
\column{0.48\textwidth}
\begin{itemize}
  \item \textbf{Unchecked exception}: subclass of \texttt{RuntimeException} (e.g., \texttt{NullPointerException}).
  \item \textbf{Stack trace}: call history showing where an exception occurred.
  \item \textbf{Handler}: catch block that responds to an exception.
\end{itemize}
\end{columns}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Study Flow for This Lecture}
\begin{enumerate}[<+->]
  \item Refresh the prerequisite bridge before reading new ideas in isolation.
  \item Study the main build in this order: \textit{Exceptions vs Errors} $\rightarrow$ \textit{What happens when an exception occurs? (stack unwinding)} $\rightarrow$ \textit{How to read a stack trace (very important skill)} $\rightarrow$ \textit{Checked vs unchecked exceptions (when to use which?)}.
  \item Trace the worked example and connect each line back to one concept frame.
  \item Attempt the mini exercises before moving to the recap and exit check.
\end{enumerate}
  \slidesources{\OOPSUnitIIISources}
\end{frame}


\section{Key Topics}
\label{sec:concepts}

\begin{frame}{Unit 03: Nested Classes, Exceptions, Multithreading \& IO Streams}
  \begin{itemize}[<+->]
    \item Lecture focus: Exception Handling (Core)
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Today: Roadmap}
  \begin{itemize}[<+->]
    \item Exceptions vs errors; checked vs unchecked
    \item `try/catch/finally`, multiple catch; `throw` vs `throws`
    \item Exception handlers + basic custom exception design for validation
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Exception vs Error}
  \begin{itemize}[<+->]
    \item Exceptions: conditions you can handle (invalid input, missing file).
    \item Errors: serious JVM/system problems (usually not recoverable).
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Checked vs Unchecked}
  \begin{itemize}[<+->]
    \item Checked: must be caught or declared (\texttt{throws}) to compile.
    \item Unchecked: \texttt{RuntimeException} (compiler does not force handling).
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{\texttt{try/catch/finally} (pattern)}
  \begin{itemize}[<+->]
    \item Put risky code in \texttt{try}.
    \item Catch specific exceptions first.
    \item Use \texttt{finally} for cleanup (runs even if exception occurs).
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{\texttt{throw} vs \texttt{throws}}
  \begin{itemize}[<+->]
    \item \texttt{throw}: create and throw an exception now.
    \item \texttt{throws}: declare that a method may throw an exception.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Example: \texttt{throw} (happens now)}
\begin{lstlisting}
static int divide(int a, int b) {
    if (b == 0) {
        throw new IllegalArgumentException("b must not be 0");
    }
    return a / b;
}
\end{lstlisting}
  \begin{itemize}[<+->]
    \item Thrown at runtime when \texttt{b == 0}.
    \item Execution jumps to a matching \texttt{catch}.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Example: \texttt{throws} (declared in signature)}
\begin{lstlisting}
static void pauseOneSecond() throws InterruptedException {
    Thread.sleep(1000); // checked exception
}
\end{lstlisting}
  \begin{itemize}[<+->]
    \item \texttt{throws} tells the caller: catch or declare.
    \item Checked exceptions are enforced by the compiler.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Caller must handle (try/catch) or declare (throws)}
\begin{lstlisting}
public static void main(String[] args) {
    try {
        pauseOneSecond();
    } catch (InterruptedException ex) {
        System.out.println("Interrupted");
    }
}
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Checkpoint and Reflection}
\begin{block}{Reflection prompt}
Where would \textit{Exception Handling (Core)} matter most in Campus Resource Manager, and which design choice would you justify using this lecture's ideas?
\end{block}
\begin{block}{Quick self-check}
Which part needs one more example before you move on: \textit{Exceptions vs Errors}, \textit{What happens when an exception occurs? (stack unwinding)}, the worked example, or the recap?
\end{block}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\section{Demo / Example}
\label{sec:demo}

\begin{frame}[fragile,shrink=8]{Full example (1/2): \texttt{try} block}
\begin{lstlisting}
import java.util.Scanner;

public class DivideDemo {
    public static void main(String[] args) {
        Scanner sc = new Scanner(System.in);

        try {
            System.out.print("Enter a: ");
            int a = Integer.parseInt(sc.nextLine());
            System.out.print("Enter b: ");
            int b = Integer.parseInt(sc.nextLine());

            int q = a / b;
            System.out.println("a/b = " + q);
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Full example (2/2): \texttt{catch} + \texttt{finally}}
\begin{lstlisting}
        } catch (NumberFormatException ex) {
            System.out.println("Please enter valid integers.");
        } catch (ArithmeticException ex) {
            System.out.println("Cannot divide by zero.");
        } finally {
            sc.close();
            System.out.println("Cleanup done (finally).");
        }
    }
}
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{How the example behaves}
  \begin{itemize}[<+->]
    \item Valid input: prints quotient, then runs \texttt{finally}.
    \item Invalid integer input: \texttt{NumberFormatException} catch runs, then \texttt{finally}.
    \item Divide by zero: \texttt{ArithmeticException} catch runs, then \texttt{finally}.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Execution sequence (try/catch/finally)}
  \begin{enumerate}[<+->]
    \item Enter \texttt{try} and execute statements top-to-bottom.
    \item If no exception: skip all \texttt{catch} blocks.
    \item If an exception occurs: jump to the \textbf{first matching} \texttt{catch}.
    \item After \texttt{try} or \texttt{catch}: always execute \texttt{finally} (cleanup).
    \item Continue with the next statement after the block (or propagate if not handled).
  \end{enumerate}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile,shrink=8]{Example: custom checked exception}
\begin{lstlisting}
class InsufficientFundsException extends Exception {
    InsufficientFundsException(String msg) { super(msg); }
}

class BankAccount {
    private double balance;
    BankAccount(double opening) { balance = opening; }
    void withdraw(double amount) throws InsufficientFundsException {
        if (amount > balance) throw new InsufficientFundsException("need " + amount);
        balance -= amount;
    }
}
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}[fragile]{Handling it}
\begin{lstlisting}
public class Demo {
    public static void main(String[] args) {
        BankAccount a = new BankAccount(100);
        try {
            a.withdraw(150);
        } catch (InsufficientFundsException e) {
            System.out.println("Withdraw failed: " + e.getMessage());
        }
    }
}
\end{lstlisting}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Mini Exercises (2--3 mins)}
  \begin{enumerate}[<+->]
    \item Give one checked and one unchecked exception example.
    \item When does \texttt{finally} run?
    \item Difference between \texttt{throw} and \texttt{throws}?
  \end{enumerate}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Watch-outs}
\begin{itemize}[<+->]
  \item Using exceptions for normal control flow.
  \item Swallowing an exception silently.
  \item Showing raw technical exception details to end users.
\end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\section{Summary}
\label{sec:summary}

\begin{frame}{Key Takeaways}
  \begin{itemize}[<+->]
    \item Exceptions enable safe error handling without crashing.
    \item Catch specific exceptions; use \texttt{finally} for cleanup.
    \item Custom exceptions make domain errors clear.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Checkpoint Questions}
  \begin{enumerate}[<+->]
    \item Why catch specific exceptions (not \texttt{Exception})?
    \item What does ``checked exception'' mean?
    \item When do we use \texttt{throws} in a method signature?
  \end{enumerate}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Next Study Steps}
\begin{itemize}[<+->]
  \item Revisit the recall references if any older idea still feels weak.
  \item Rework the lecture example without looking at the notes first.
  \item Attempt the self-study practice and then answer the exit questions in full sentences.
  \item Apply the idea once inside Campus Resource Manager to make the concept stick.
\end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\begin{frame}{Next Lecture Preview}
  \textbf{Next:} Multithreading + Synchronization
  \vspace{0.6em}
  \begin{itemize}[<+->]
    \item We will create threads using \texttt{Thread} and \texttt{Runnable} and see race conditions.
  \end{itemize}
  \slidesources{\OOPSUnitIIISources}
\end{frame}

\end{document}
