\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming Lab (CSEG1044)\\Experiment-15\&16: Lambdas + Swing GUI (Student Handout --- Before Submission)}
\author{Tofik Ali}
\date{\today}

\begin{document}
\maketitle
\tableofcontents

\section{About this handout}
This handout specifies the required programs for lambda expressions and Swing GUI.
Do the implementation yourself and include outputs/screenshots.

\section{Experiment 15 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To use lambda expressions and functional interfaces.
  \item \textbf{Problem Statement:} Implement lambda expressions in Java programs.
  \item \textbf{Expected Outcome:} Students will understand functional programming.
\end{itemize}

\section{Experiment 16 (Syllabus)}
\begin{itemize}[leftmargin=*]
  \item \textbf{Objective:} To design GUI applications using Swing.
  \item \textbf{Problem Statement:} Develop GUI applications using Swing components.
  \item \textbf{Expected Outcome:} Students will create interactive applications.
\end{itemize}

\section{What you must do (minimum requirements)}

\subsection{Task A --- Lambda practice (Exp-15)}
Create \texttt{LambdaDemo.java}:
\begin{enumerate}[leftmargin=*]
  \item Define a custom \texttt{@FunctionalInterface} (example: \texttt{IntOp}) with method \texttt{apply(int a, int b)}.
  \item Create lambdas:
  \begin{itemize}[leftmargin=*]
    \item addition
    \item max/min
  \end{itemize}
  \item Create a \texttt{Runnable} using lambda and run it in a thread.
  \item Sort a string array using a comparator lambda (case-insensitive).
\end{enumerate}

\subsection{Task B --- Swing UI with events (Exp-16)}
Create \texttt{SwingDemo.java}:
\begin{enumerate}[leftmargin=*]
  \item Create a window with:
  \begin{itemize}[leftmargin=*]
    \item text field for name
    \item button ``Greet''
    \item label for output
    \item button ``Reset''
  \end{itemize}
  \item On ``Greet'' click:
  \begin{itemize}[leftmargin=*]
    \item increment a click counter
    \item update label: \texttt{Hello <name> (Clicks: X)}
  \end{itemize}
  \item On ``Reset'' click: set counter to 0 and clear the text field.
  \item Use \texttt{SwingUtilities.invokeLater(...)} (EDT best practice).
\end{enumerate}

\section{Hints (read only if stuck)}
\begin{itemize}[leftmargin=*]
  \item A lambda can be used only where a functional interface is expected.
  \item In Swing, create and update UI on the Event Dispatch Thread (EDT).
  \item Use an action listener to handle button clicks.
\end{itemize}

\section{What to submit (MANDATORY)}
Submit a single zipped folder containing:
\begin{itemize}[leftmargin=*]
  \item \texttt{LambdaDemo.java}
  \item \texttt{SwingDemo.java}
  \item sample output for \texttt{LambdaDemo.java}
  \item screenshot of the Swing UI running (or short recording)
  \item observations (5--10 lines): where lambdas helped; what EDT is
\end{itemize}

\notesources{\OOPSUnitIVSources}
\end{document}
