\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming (CSEG1044)\\Unit 05 -- Lecture 02 Notes\\Set/SortedSet + Map/SortedMap}
\begin{document}
\maketitle
\tableofcontents

\section{Lecture Snapshot}
\begin{itemize}
  \item \textbf{Unit focus:} Collections Framework and Wrapper Classes
  \item \textbf{Main new ideas:} Set vs List; HashSet vs TreeSet, Equality and hashing basics (must know), Map basics; HashMap vs TreeMap
  \item \textbf{Course thread:} Use Set/SortedSet + Map/SortedMap to manage in-memory project data more cleanly and predictably inside Campus Resource Manager.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Lecture Overview}
This lecture focuses on \textbf{uniqueness} (Set) and \textbf{key-value lookup} (Map).
These structures are essential for problems like:
\begin{itemize}
  \item removing duplicates,
  \item membership checks (``is this item present?''),
  \item counting and grouping (frequency tables),
  \item fast search by key (roll number $\rightarrow$ student details).
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Prerequisite Bridge}
Before reading the new material in depth, do a fast recall of the older ideas listed below.
\begin{itemize}
  \item \textbf{List basics}: Sets and maps extend the collection story by optimizing for uniqueness and key-based lookup. Main reference: \texttt{Unit 05 / Lecture 01 slides/notes}.
  \item \textbf{Class design}: Set and map correctness depends on stable equality logic for stored objects. Main reference: \texttt{Unit 02 / Lecture 01 slides/notes}.
\end{itemize}
This lecture only revisits those ideas briefly so that the main explanation can stay focused on the new concept sequence.
\notesources{\OOPSUnitVSources}

\section{Learning Outcomes}
After this lecture, you should be able to:
\begin{itemize}
  \item Choose Set vs List and \texttt{HashSet} vs \texttt{TreeSet}.
  \item Explain \texttt{equals}/\texttt{hashCode} and common mistakes.
  \item Use \texttt{HashMap}/\texttt{TreeMap} for counting and sorted keys.
\end{itemize}

\section{Suggested Reading Order}
\begin{enumerate}
  \item Read the \textbf{Prerequisite Bridge} first and confirm each recalled idea.
  \item Study the central explanation in this order: \textit{Set vs List; HashSet vs TreeSet} $\rightarrow$ \textit{Equality and hashing basics (must know)} $\rightarrow$ \textit{Map basics; HashMap vs TreeMap}.
  \item Trace the worked example line by line before checking short answers.
  \item Attempt the practice section without looking at the solution immediately.
  \item Finish with the exit questions and further-study suggestions to confirm retention.
\end{enumerate}
\notesources{\OOPSUnitVSources}

\section{Key Terms (Quick)}
\begin{itemize}
  \item \textbf{Set}: collection with no duplicates.
  \item \textbf{SortedSet}: set that maintains sorted order (e.g., \texttt{TreeSet}).
  \item \textbf{Hashing}: converting an object to an int hash value (\texttt{hashCode}).
  \item \textbf{Map}: key-value association.
  \item \textbf{SortedMap}: map with sorted keys (e.g., \texttt{TreeMap}).
  \item \textbf{Contract}: if \texttt{a.equals(b)} then \texttt{a.hashCode() == b.hashCode()} must hold.
\end{itemize}

\section{Detailed Notes}
\subsection{Set vs List; HashSet vs TreeSet}
\textbf{List:}
\begin{itemize}
  \item ordered, duplicates allowed, index-based access.
\end{itemize}
\textbf{Set:}
\begin{itemize}
  \item no duplicates, typically no index access.
  \item ideal for membership: ``contains?'' and deduplication.
\end{itemize}

\textbf{HashSet:}
\begin{itemize}
  \item uses hashing; order is not guaranteed.
  \item fast average-case add/contains/remove.
\end{itemize}

\textbf{TreeSet:}
\begin{itemize}
  \item keeps elements sorted (needs \texttt{Comparable} or a \texttt{Comparator}).
  \item slightly slower but provides sorted iteration.
\end{itemize}

\subsection{Equality and hashing basics (must know)}
Hash-based collections (\texttt{HashSet}, \texttt{HashMap}) use:
\begin{itemize}
  \item \texttt{hashCode()} to choose a bucket,
  \item \texttt{equals()} to confirm identity inside that bucket.
\end{itemize}

\textbf{Contract:}
\begin{itemize}
  \item If two objects are equal (\texttt{equals} returns true), their hash codes must be equal.
  \item If you override \texttt{equals}, you should also override \texttt{hashCode}.
\end{itemize}

\textbf{Common mistakes:}
\begin{itemize}
  \item overriding only one of \texttt{equals}/\texttt{hashCode},
  \item using mutable fields in keys (changing key after insertion breaks map/set behavior),
  \item using \texttt{==} instead of \texttt{equals} for strings.
\end{itemize}

\subsection{Map basics; HashMap vs TreeMap}
\textbf{Map} stores pairs: key $\rightarrow$ value.
Common operations:
\begin{itemize}
  \item \texttt{put(k, v)}, \texttt{get(k)}, \texttt{containsKey(k)}, \texttt{remove(k)}.
\end{itemize}

\textbf{HashMap:} no ordering, fast average-case lookup.
\textbf{TreeMap:} keys sorted, good when you need ordered iteration or range queries.
\notesources{\OOPSUnitVSources}

\section{Running Project Connection}
Use Set/SortedSet + Map/SortedMap to manage in-memory project data more cleanly and predictably inside Campus Resource Manager.
\begin{itemize}
  \item Use Campus Resource Manager as the running case study, or map the same idea to your own project domain if you are using another example.
  \item Ask where today's idea should live: model, service, DAO, UI, or team workflow.
  \item Use the lecture example as a smaller version of the same design choice you will make in the capstone.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Common Mistakes and Confusions}
\begin{itemize}
  \item Forgetting the equals/hashCode contract and blaming the collection.
  \item Expecting HashSet or HashMap to preserve sorted order.
  \item Using a map or set before confirming the problem truly needs keys or uniqueness.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Worked Example: set + map counting}
\begin{lstlisting}
import java.util.*;

public class Demo {
    public static void main(String[] args) {
        // Dedup using Set
        List<String> names = Arrays.asList("Asha", "Ravi", "Asha", "Neha");
        Set<String> unique = new HashSet<>(names);
        System.out.println("Unique: " + unique);

        // Counting with Map (frequency table)
        String[] words = {"a", "b", "a", "c", "b", "a"};
        Map<String, Integer> freq = new HashMap<>();
        for (String w : words) {
            freq.put(w, freq.getOrDefault(w, 0) + 1);
        }
        System.out.println("Freq: " + freq);

        // Sorted keys using TreeMap
        Map<String, Integer> sorted = new TreeMap<>(freq);
        System.out.println("Sorted freq: " + sorted);
    }
}
\end{lstlisting}

\section{Demo / Observation Task}
Use this block as a short reproducible demo or observation task.
\begin{itemize}
  \item Reproduce the lecture's main example around \textit{Set vs List; HashSet vs TreeSet} once without looking at the solution first.
  \item Change one input, rule, or design choice in Campus Resource Manager and predict the result before checking it.
  \item Record one success case, one edge case, and one failure case so the idea becomes observable instead of purely theoretical.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Practice Check (with brief solutions)}
\begin{enumerate}
  \item When should you prefer a Set over a List?\par
  \textbf{Answer:} when duplicates are not allowed or you need fast membership checks.

  \item Which collection keeps elements sorted: \texttt{HashSet} or \texttt{TreeSet}?\par
  \textbf{Answer:} \texttt{TreeSet}.

  \item If you override \texttt{equals}, what else should you override?\par
  \textbf{Answer:} \texttt{hashCode}.
\end{enumerate}

\section{Self-Study Practice (no solutions)}
\begin{enumerate}
  \item Count character frequency in a string using \texttt{Map<Character, Integer>}.
  \item Store student details in a map: rollNo $\rightarrow$ Student object.
  \item Create a \texttt{TreeSet} of integers and print them in sorted order.
\end{enumerate}

\section{Further Study}
\begin{itemize}
  \item Revisit the prerequisite references if any recall bullet still feels weak.
  \item Rework the lecture example with one changed assumption, input, or design choice.
  \item Read the textbook and documentation references listed in the source line for a second explanation of the same topic.
  \item Apply the same idea once inside Campus Resource Manager so that the concept moves from theory to design practice.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Exit Questions (with sample answers)}
\textbf{Q1.} Why are \texttt{equals} and \texttt{hashCode} important for \texttt{HashSet}?\par
\textbf{Sample answer:} \texttt{HashSet} uses \texttt{hashCode} to locate a bucket and \texttt{equals} to check if an element already exists. Incorrect implementations can cause duplicates or missing lookups.

\vspace{0.6em}
\textbf{Q2.} When do you choose \texttt{TreeMap} over \texttt{HashMap}?\par
\textbf{Sample answer:} Choose \texttt{TreeMap} when you need keys in sorted order or want ordered iteration/range queries. Use \texttt{HashMap} when ordering is not needed and you want faster average lookup.

\vspace{0.6em}
\textbf{Q3.} When should TreeSet or TreeMap be preferred over hash-based versions?\par
\textbf{Sample answer:} When the requirement needs sorted traversal, natural ordering, or range-style behavior.

\end{document}
