\documentclass{beamer}

\input{../../../_shared/beamer-common.tex}

\title[OOPS with Java]{Object Oriented Programming (CSEG1044)}
\subtitle{Unit 05 -- Lecture 02: Set/SortedSet + Map/SortedMap}
\begin{document}

\begin{frame}[fragile]
  \titlepage
  \vspace{0.5em}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Quick Links}
  \centering
  \hyperlink{sec:overview}{\beamerbutton{Overview}}\hspace{0.6em}
  \hyperlink{sec:concepts}{\beamerbutton{Key Topics}}\hspace{0.6em}
  \hyperlink{sec:demo}{\beamerbutton{Demo}}\hspace{0.6em}
  \hyperlink{sec:summary}{\beamerbutton{Summary}}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Agenda}
  \tableofcontents
  \slidesources{\OOPSUnitVSources}
\end{frame}

\section{Overview}
\label{sec:overview}

\begin{frame}{Lecture Snapshot}
\begin{itemize}[<+->]
  \item \textbf{Unit focus:} Collections Framework and Wrapper Classes
  \item \textbf{Course thread:} Use Set/SortedSet + Map/SortedMap to manage in-memory project data more cleanly and predictably inside Campus Resource Manager.
\end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Recall Before You Start}
\begin{itemize}[<+->]
  \item \textbf{List basics}: Sets and maps extend the collection story by optimizing for uniqueness and key-based lookup. Main reference: \texttt{Unit 05 / Lecture 01 slides/notes}.
  \item \textbf{Class design}: Set and map correctness depends on stable equality logic for stored objects. Main reference: \texttt{Unit 02 / Lecture 01 slides/notes}.
\end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}


\begin{frame}{Learning Outcomes}
  \begin{itemize}[<+->]
    \item Choose Set vs List and \texttt{HashSet} vs \texttt{TreeSet}.
    \item Explain \texttt{equals}/\texttt{hashCode} and common mistakes.
    \item Use \texttt{HashMap}/\texttt{TreeMap} for counting and sorted keys.
  \end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Key Terms to Know}
\small
\begin{columns}[t]
\column{0.48\textwidth}
\begin{itemize}
  \item \textbf{Set}: collection with no duplicates.
  \item \textbf{SortedSet}: set that maintains sorted order (e.g., \texttt{TreeSet}).
  \item \textbf{Hashing}: converting an object to an int hash value (\texttt{hashCode}).
\end{itemize}
\column{0.48\textwidth}
\begin{itemize}
  \item \textbf{Map}: key-value association.
  \item \textbf{SortedMap}: map with sorted keys (e.g., \texttt{TreeMap}).
  \item \textbf{Contract}: if \texttt{a.equals(b)} then \texttt{a.hashCode() == b.hashCode()} must hold.
\end{itemize}
\end{columns}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Study Flow for This Lecture}
\begin{enumerate}[<+->]
  \item Refresh the prerequisite bridge before reading new ideas in isolation.
  \item Study the main build in this order: \textit{Set vs List; HashSet vs TreeSet} $\rightarrow$ \textit{Equality and hashing basics (must know)} $\rightarrow$ \textit{Map basics; HashMap vs TreeMap}.
  \item Trace the worked example and connect each line back to one concept frame.
  \item Attempt the mini exercises before moving to the recap and exit check.
\end{enumerate}
  \slidesources{\OOPSUnitVSources}
\end{frame}


\section{Key Topics}
\label{sec:concepts}

\begin{frame}{Unit 05: Collections and Wrapper Class}
  \begin{itemize}[<+->]
    \item Lecture focus: Set/SortedSet + Map/SortedMap
  \end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Today: Roadmap}
  \begin{itemize}[<+->]
    \item Set vs list; `HashSet` vs `TreeSet`
    \item Equality and hashing basics (`equals`/`hashCode`) and common mistakes
    \item Map basics; `HashMap` vs `TreeMap`; common use-cases (counting/grouping)
  \end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Set vs List}
  \begin{itemize}[<+->]
    \item List: ordered, duplicates allowed.
    \item Set: no duplicates; great for membership and deduplication.
  \end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{HashSet vs TreeSet}
  \begin{itemize}[<+->]
    \item \texttt{HashSet}: no guaranteed order; fast average operations.
    \item \texttt{TreeSet}: sorted order; needs \texttt{Comparable} or \texttt{Comparator}.
  \end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Why \texttt{equals} and \texttt{hashCode} matter}
  \begin{itemize}[<+->]
    \item Hash-based collections use \texttt{hashCode} to find a bucket.
    \item \texttt{equals} confirms the match inside the bucket.
    \item Contract: if \texttt{a.equals(b)} then \texttt{a.hashCode()==b.hashCode()}.
  \end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Map basics}
  \begin{itemize}[<+->]
    \item Map = key $\rightarrow$ value.
    \item Use cases: counting, grouping, lookups by id.
    \item \texttt{HashMap}: fast average; \texttt{TreeMap}: keys sorted.
  \end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Checkpoint and Reflection}
\begin{block}{Reflection prompt}
Where would \textit{Set/SortedSet + Map/SortedMap} matter most in Campus Resource Manager, and which design choice would you justify using this lecture's ideas?
\end{block}
\begin{block}{Quick self-check}
Which part needs one more example before you move on: \textit{Set vs List; HashSet vs TreeSet}, \textit{Equality and hashing basics (must know)}, the worked example, or the recap?
\end{block}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\section{Demo / Example}
\label{sec:demo}

\begin{frame}[fragile,shrink=8]{Example: Set + Map counting}
\begin{lstlisting}
import java.util.*;
public class Demo {
    public static void main(String[] args) {
        List<String> names = Arrays.asList("Asha", "Ravi", "Asha");
        Set<String> unique = new HashSet<>(names);
        System.out.println(unique);

        String[] words = {"a","b","a","c","b","a"};
        Map<String,Integer> freq = new HashMap<>();
        for (String w : words) freq.put(w, freq.getOrDefault(w,0) + 1);
        System.out.println(freq);

        Map<String,Integer> sorted = new TreeMap<>(freq);
        System.out.println(sorted);
    }
}
\end{lstlisting}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Mini Exercises (2--3 mins)}
  \begin{enumerate}[<+->]
    \item When do you prefer Set over List?
    \item Which keeps sorted order: \texttt{HashSet} or \texttt{TreeSet}?
    \item If you override \texttt{equals}, what else must you override?
  \end{enumerate}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Watch-outs}
\begin{itemize}[<+->]
  \item Forgetting the equals/hashCode contract and blaming the collection.
  \item Expecting HashSet or HashMap to preserve sorted order.
  \item Using a map or set before confirming the problem truly needs keys or uniqueness.
\end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\section{Summary}
\label{sec:summary}

\begin{frame}{Key Takeaways}
  \begin{itemize}[<+->]
    \item Set enforces uniqueness; TreeSet provides sorted order.
    \item Hashing relies on correct \texttt{equals}/\texttt{hashCode}.
    \item Map enables fast key-based lookup and counting/grouping.
  \end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Checkpoint Questions}
  \begin{enumerate}[<+->]
    \item Why do Hash-based collections need \texttt{hashCode} and \texttt{equals}?
    \item When do you choose \texttt{TreeMap} over \texttt{HashMap}?
    \item Give one example of counting/grouping using a map.
  \end{enumerate}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Next Study Steps}
\begin{itemize}[<+->]
  \item Revisit the recall references if any older idea still feels weak.
  \item Rework the lecture example without looking at the notes first.
  \item Attempt the self-study practice and then answer the exit questions in full sentences.
  \item Apply the idea once inside Campus Resource Manager to make the concept stick.
\end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\begin{frame}{Next Lecture Preview}
  \textbf{Next:} Wrapped Collections + \texttt{Collections} utilities + Wrapper classes
  \vspace{0.6em}
  \begin{itemize}[<+->]
    \item Utility methods and common pitfalls with autoboxing.
  \end{itemize}
  \slidesources{\OOPSUnitVSources}
\end{frame}

\end{document}
