\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming (CSEG1044)\\Unit 04 -- Lecture 05 Notes\\Swing Event Handling + JDBC (Persistence)}
\begin{document}
\maketitle
\tableofcontents

\section{Lecture Snapshot}
\begin{itemize}
  \item \textbf{Unit focus:} Generics, Lambdas, Swing, and JDBC
  \item \textbf{Main new ideas:} Swing event model (listeners), JDBC core flow (pattern you should memorize), \texttt{Statement} vs \texttt{PreparedStatement} vs \texttt{CallableStatement}
  \item \textbf{Course thread:} Use Swing Event Handling + JDBC (Persistence) to move Campus Resource Manager toward a real desktop application with UI, callbacks, and persistence.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Lecture Overview}
In the last lecture we built a Swing UI skeleton.
Now we make it interactive and persistent:
\begin{itemize}
  \item \textbf{Events}: a button click triggers code (save/clear/exit).
  \item \textbf{Persistence}: data is stored in a database using JDBC so it is not lost after closing the app.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Prerequisite Bridge}
Before reading the new material in depth, do a fast recall of the older ideas listed below.
\begin{itemize}
  \item \textbf{Swing foundations}: You already know the component structure; now user actions and persistence are attached to that UI. Main reference: \texttt{Unit 04 / Lecture 04 slides/notes}.
  \item \textbf{Failure handling}: JDBC and event code both need disciplined exception handling and user messages. Main reference: \texttt{Unit 03 / Lecture 02 slides/notes}.
\end{itemize}
This lecture only revisits those ideas briefly so that the main explanation can stay focused on the new concept sequence.
\notesources{\OOPSUnitIVSources}

\section{Learning Outcomes}
After this lecture, you should be able to:
\begin{itemize}
  \item Connect Swing UI controls to actions using listeners.
  \item Explain JDBC flow and choose between \texttt{Statement} and \texttt{PreparedStatement}.
  \item Implement basic CRUD and read results using \texttt{ResultSet}.
\end{itemize}

\section{Suggested Reading Order}
\begin{enumerate}
  \item Read the \textbf{Prerequisite Bridge} first and confirm each recalled idea.
  \item Study the central explanation in this order: \textit{Swing event model (listeners)} $\rightarrow$ \textit{JDBC core flow (pattern you should memorize)} $\rightarrow$ \textit{\texttt{Statement} vs \texttt{PreparedStatement} vs \texttt{CallableStatement}} $\rightarrow$ \textit{\texttt{ResultSet} traversal} $\rightarrow$ \textit{Persistent data design (UI vs DAO)}.
  \item Trace the worked example line by line before checking short answers.
  \item Attempt the practice section without looking at the solution immediately.
  \item Finish with the exit questions and further-study suggestions to confirm retention.
\end{enumerate}
\notesources{\OOPSUnitIVSources}

\section{Key Terms (Quick)}
\begin{itemize}
  \item \textbf{Event source}: UI component that triggers an event (button, menu item).
  \item \textbf{Event object}: information about the event (e.g., \texttt{ActionEvent}).
  \item \textbf{Listener}: object/function that handles events (e.g., \texttt{ActionListener}).
  \item \textbf{EDT}: Event Dispatch Thread (Swing UI thread).
  \item \textbf{JDBC}: Java Database Connectivity API.
  \item \textbf{Connection}: active DB connection.
  \item \textbf{PreparedStatement}: parameterized SQL statement (safer and common).
  \item \textbf{ResultSet}: table-like cursor over query results.
  \item \textbf{CRUD}: Create, Read, Update, Delete.
  \item \textbf{DAO}: Data Access Object (separate DB code from UI).
\end{itemize}

\section{Detailed Notes}
\subsection{Swing event model (listeners)}
Swing uses a listener-based model:
\begin{itemize}
  \item You create a component (e.g., \texttt{JButton saveBtn}).
  \item You register a listener: \texttt{saveBtn.addActionListener(...)}.
  \item When user clicks, Swing calls your listener method on the EDT.
\end{itemize}

\textbf{Common controls and events:}
\begin{itemize}
  \item \texttt{JButton} $\rightarrow$ \texttt{ActionListener}
  \item \texttt{JTextField} $\rightarrow$ \texttt{ActionListener} (press Enter), \texttt{DocumentListener} (text changes)
  \item \texttt{JTable} $\rightarrow$ selection listeners (advanced)
\end{itemize}

\paragraph{Important rule.} Do not run slow DB operations on the EDT; it can freeze the UI.
In a basic course project, keep operations small or use a background thread/SwingWorker (advanced).

\subsection{JDBC core flow (pattern you should memorize)}
\begin{enumerate}
  \item Load/ensure driver is available (often automatic in modern JDBC).
  \item Get a connection: \texttt{DriverManager.getConnection(url, user, pass)}.
  \item Create a statement (usually \texttt{PreparedStatement}).
  \item Execute:
  \begin{itemize}
    \item \texttt{executeUpdate()} for INSERT/UPDATE/DELETE
    \item \texttt{executeQuery()} for SELECT
  \end{itemize}
  \item Read results using \texttt{ResultSet} (cursor).
  \item Close resources (use \texttt{try-with-resources}).
\end{enumerate}

\subsection{\texttt{Statement} vs \texttt{PreparedStatement} vs \texttt{CallableStatement}}
\begin{itemize}
  \item \texttt{Statement}: simplest; builds SQL as a string (risk of SQL injection; avoid for user input).
  \item \texttt{PreparedStatement}: parameterized with \texttt{?}; safer and faster for repeated queries.
  \item \texttt{CallableStatement}: calls stored procedures (depends on DB).
\end{itemize}

\subsection{\texttt{ResultSet} traversal}
\begin{itemize}
  \item \texttt{while (rs.next()) \{ ... \}} moves row by row.
  \item Read columns by label or index: \texttt{rs.getInt("id")}, \texttt{rs.getString("name")}.
  \item Map each row to an object (e.g., \texttt{Resource}).
\end{itemize}

\subsection{Persistent data design (UI vs DAO)}
Avoid writing SQL inside button click code.
Preferred structure:
\begin{itemize}
  \item UI layer: reads fields, validates, shows messages.
  \item DAO layer: contains SQL and JDBC code.
  \item Model layer: \texttt{Resource} class that represents one row/entity.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Running Project Connection}
Use Swing Event Handling + JDBC (Persistence) to move Campus Resource Manager toward a real desktop application with UI, callbacks, and persistence.
\begin{itemize}
  \item Use Campus Resource Manager as the running case study, or map the same idea to your own project domain if you are using another example.
  \item Ask where today's idea should live: model, service, DAO, UI, or team workflow.
  \item Use the lecture example as a smaller version of the same design choice you will make in the capstone.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Common Mistakes and Confusions}
\begin{itemize}
  \item Building SQL strings directly from user input.
  \item Running slow database work on the Event Dispatch Thread.
  \item Mixing UI controls, validation, and DAO logic in one method.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Worked Example: Save a resource from Swing to DB (skeleton)}
\subsection*{1) Model class}
\begin{lstlisting}
public class Resource {
    private final String name;
    private final String type;
    private final int capacity;

    public Resource(String name, String type, int capacity) {
        this.name = name;
        this.type = type;
        this.capacity = capacity;
    }

    public String getName() { return name; }
    public String getType() { return type; }
    public int getCapacity() { return capacity; }
}
\end{lstlisting}

\subsection*{2) DAO (insert + select)}
\begin{lstlisting}
import java.sql.*;
import java.util.ArrayList;
import java.util.List;

public class ResourceDao {
    private final String url = "jdbc:mysql://localhost:3306/crm"; // example
    private final String user = "root";
    private final String pass = "password";

    public void insert(Resource r) throws SQLException {
        String sql = "INSERT INTO resources(name, type, capacity) VALUES(?,?,?)";
        try (Connection con = DriverManager.getConnection(url, user, pass);
             PreparedStatement ps = con.prepareStatement(sql)) {
            ps.setString(1, r.getName());
            ps.setString(2, r.getType());
            ps.setInt(3, r.getCapacity());
            ps.executeUpdate();
        }
    }

    public List<Resource> findAll() throws SQLException {
        String sql = "SELECT name, type, capacity FROM resources";
        List<Resource> out = new ArrayList<>();
        try (Connection con = DriverManager.getConnection(url, user, pass);
             PreparedStatement ps = con.prepareStatement(sql);
             ResultSet rs = ps.executeQuery()) {
            while (rs.next()) {
                out.add(new Resource(
                    rs.getString("name"),
                    rs.getString("type"),
                    rs.getInt("capacity")
                ));
            }
        }
        return out;
    }
}
\end{lstlisting}

\subsection*{3) Event handler (button click)}
\begin{lstlisting}
saveBtn.addActionListener(e -> {
    String name = nameField.getText().trim();
    String type = typeField.getText().trim();
    String capText = capacityField.getText().trim();

    if (name.isEmpty() || type.isEmpty() || capText.isEmpty()) {
        JOptionPane.showMessageDialog(f, "All fields required");
        return;
    }

    int cap;
    try {
        cap = Integer.parseInt(capText);
        if (cap <= 0) throw new NumberFormatException();
    } catch (NumberFormatException ex) {
        JOptionPane.showMessageDialog(f, "Capacity must be a positive integer");
        return;
    }

    saveInBackground(name, type, cap);
});
\end{lstlisting}
The listener validates input and delegates the JDBC work. The DB call should run in a background task. \texttt{SwingWorker} is a simple option that keeps the EDT responsive.

\section{Demo / Observation Task}
Use this block as a short reproducible demo or observation task.
\begin{itemize}
  \item Reproduce the lecture's main example around \textit{Swing event model (listeners)} once without looking at the solution first.
  \item Change one input, rule, or design choice in Campus Resource Manager and predict the result before checking it.
  \item Record one success case, one edge case, and one failure case so the idea becomes observable instead of purely theoretical.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Practice Check (with brief solutions)}
\begin{enumerate}
  \item Which is safer for user input: \texttt{Statement} or \texttt{PreparedStatement}?\par
  \textbf{Answer:} \texttt{PreparedStatement}.

  \item Why should heavy DB operations not run on the EDT?\par
  \textbf{Answer:} the UI can freeze because Swing processes events on the EDT.

  \item What method is used for SELECT queries: \texttt{executeQuery()} or \texttt{executeUpdate()}?\par
  \textbf{Answer:} \texttt{executeQuery()}.
\end{enumerate}

\section{Self-Study Practice (no solutions)}
\begin{enumerate}
  \item Add an \texttt{updateCapacity(name, newCapacity)} method in \texttt{ResourceDao}.
  \item Add a ``Load All'' button that displays all resources in a text area/table.
  \item Replace DB credentials with a config file and load them at startup (optional).
\end{enumerate}

\section{Lab Alignment (recommended)}
\begin{itemize}
  \item Exp-16: GUI using Swing
  \item Exp-17: Database connectivity using JDBC
\end{itemize}

\section{Further Study}
\begin{itemize}
  \item Revisit the prerequisite references if any recall bullet still feels weak.
  \item Rework the lecture example with one changed assumption, input, or design choice.
  \item Read the textbook and documentation references listed in the source line for a second explanation of the same topic.
  \item Apply the same idea once inside Campus Resource Manager so that the concept moves from theory to design practice.
\end{itemize}
\notesources{\OOPSUnitIVSources}

\section{Exit Questions (with sample answers)}
\textbf{Q1.} Why is \texttt{PreparedStatement} preferred?\par
\textbf{Sample answer:} It allows parameterized queries, prevents SQL injection for user input, and can be more efficient for repeated execution.

\vspace{0.6em}
\textbf{Q2.} What is the role of a DAO?\par
\textbf{Sample answer:} A DAO separates database code from UI/business logic, making the application easier to maintain, test, and change.

\vspace{0.6em}
\textbf{Q3.} Why is PreparedStatement the default choice for CRUD forms?\par
\textbf{Sample answer:} It supports parameterized queries and makes database operations safer and cleaner.

\end{document}
