\documentclass[11pt]{article}

\input{../../../_shared/notes-common.tex}

\title{Object Oriented Programming (CSEG1044)\\Unit 05 -- Lecture 01 Notes\\Collections Framework Overview + Iteration + List}
\begin{document}
\maketitle
\tableofcontents

\section{Lecture Snapshot}
\begin{itemize}
  \item \textbf{Unit focus:} Collections Framework and Wrapper Classes
  \item \textbf{Main new ideas:} Core interfaces: \texttt{Iterable}, \texttt{Collection}, \texttt{Iterator}, List: ArrayList vs LinkedList, Sorting: Comparable vs Comparator
  \item \textbf{Course thread:} Use Collections Framework Overview + Iteration + List to manage in-memory project data more cleanly and predictably inside Campus Resource Manager.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Lecture Overview}
The Collections Framework gives you standard, reusable data structures in Java.
Instead of creating arrays manually and writing custom logic, you use:
\begin{itemize}
  \item \texttt{List} for ordered sequences,
  \item \texttt{Set} for unique values,
  \item \texttt{Map} for key-value lookups.
\end{itemize}
This lecture focuses on \textbf{lists}, \textbf{iteration}, and \textbf{sorting}.
\notesources{\OOPSUnitVSources}

\section{Prerequisite Bridge}
Before reading the new material in depth, do a fast recall of the older ideas listed below.
\begin{itemize}
  \item \textbf{Generics}: Collections are safer and clearer once typed containers are already familiar. Main reference: \texttt{Unit 04 / Lecture 01 slides/notes}.
  \item \textbf{Comparison logic}: Sorting becomes easier once small comparison behavior is familiar. Main reference: \texttt{Unit 04 / Lecture 03 slides/notes}.
\end{itemize}
This lecture only revisits those ideas briefly so that the main explanation can stay focused on the new concept sequence.
\notesources{\OOPSUnitVSources}

\section{Learning Outcomes}
After this lecture, you should be able to:
\begin{itemize}
  \item Explain \texttt{Iterable}, \texttt{Collection}, and \texttt{Iterator}.
  \item Choose \texttt{ArrayList} vs \texttt{LinkedList} for common use cases.
  \item Sort a list using \texttt{Comparable} and \texttt{Comparator} (with lambdas).
\end{itemize}

\section{Suggested Reading Order}
\begin{enumerate}
  \item Read the \textbf{Prerequisite Bridge} first and confirm each recalled idea.
  \item Study the central explanation in this order: \textit{Core interfaces: \texttt{Iterable}, \texttt{Collection}, \texttt{Iterator}} $\rightarrow$ \textit{List: ArrayList vs LinkedList} $\rightarrow$ \textit{Sorting: Comparable vs Comparator}.
  \item Trace the worked example line by line before checking short answers.
  \item Attempt the practice section without looking at the solution immediately.
  \item Finish with the exit questions and further-study suggestions to confirm retention.
\end{enumerate}
\notesources{\OOPSUnitVSources}

\section{Key Terms (Quick)}
\begin{itemize}
  \item \textbf{Collection}: group of elements (root interface for many collections).
  \item \textbf{Iterable}: can be used in for-each loop (provides \texttt{iterator()}).
  \item \textbf{Iterator}: cursor used to traverse a collection safely.
  \item \textbf{List}: ordered collection with duplicates allowed and index-based access.
  \item \textbf{Comparable}: natural ordering (\texttt{compareTo}).
  \item \textbf{Comparator}: custom ordering (\texttt{compare}).
\end{itemize}

\section{Detailed Notes}
\subsection{Core interfaces: \texttt{Iterable}, \texttt{Collection}, \texttt{Iterator}}
\textbf{Iterable:}
\begin{itemize}
  \item Any \texttt{Iterable} can be used with the enhanced for loop.
\end{itemize}
\textbf{Iterator:}
\begin{itemize}
  \item Provides \texttt{hasNext()}, \texttt{next()}, and optional \texttt{remove()}.
  \item Use iterator when you need to remove elements while looping.
\end{itemize}

\paragraph{Common pitfall.} Removing from a list inside a for-each loop can throw \path|ConcurrentModificationException|. If removal is required, switch to an \texttt{Iterator}.

\subsection{List: ArrayList vs LinkedList}
\textbf{ArrayList:}
\begin{itemize}
  \item Backed by a dynamic array.
  \item Fast random access: \texttt{get(i)} is efficient.
  \item Good for most use cases: iterate a lot, access by index.
\end{itemize}

\textbf{LinkedList:}
\begin{itemize}
  \item Nodes with links (doubly linked list).
  \item Slower random access (must traverse nodes).
  \item Useful when you frequently add/remove at ends and use it as a queue/deque.
\end{itemize}

\paragraph{Rule of thumb.} Use \texttt{ArrayList} by default unless you have a clear reason to use \texttt{LinkedList}.

\subsection{Sorting: Comparable vs Comparator}
\textbf{Comparable:}
\begin{itemize}
  \item implemented by the class itself (natural ordering).
  \item example: sort students by roll number.
\end{itemize}

\textbf{Comparator:}
\begin{itemize}
  \item separate ordering logic, reusable and flexible.
  \item example: sort students by CGPA descending.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Running Project Connection}
Use Collections Framework Overview + Iteration + List to manage in-memory project data more cleanly and predictably inside Campus Resource Manager.
\begin{itemize}
  \item Use Campus Resource Manager as the running case study, or map the same idea to your own project domain if you are using another example.
  \item Ask where today's idea should live: model, service, DAO, UI, or team workflow.
  \item Use the lecture example as a smaller version of the same design choice you will make in the capstone.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Common Mistakes and Confusions}
\begin{itemize}
  \item Ignoring generics and using raw collections.
  \item Changing a collection unsafely while iterating.
  \item Choosing a List when the real requirement is uniqueness.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Worked Example: list + iterator + sorting}
\begin{lstlisting}
import java.util.*;

class Student {
    private final String rollNo;
    private final double cgpa;

    Student(String rollNo, double cgpa) {
        this.rollNo = rollNo;
        this.cgpa = cgpa;
    }

    public String getRollNo() { return rollNo; }
    public double getCgpa() { return cgpa; }
    public String toString() { return rollNo + ":" + cgpa; }
}

public class Demo {
    public static void main(String[] args) {
        List<Student> list = new ArrayList<>();
        list.add(new Student("21CS1002", 9.1));
        list.add(new Student("21CS1001", 8.2));
        list.add(new Student("21CS1003", 3.9));

        // Remove weak students safely using Iterator
        Iterator<Student> it = list.iterator();
        while (it.hasNext()) {
            if (it.next().getCgpa() < 4.0) it.remove();
        }

        // Sort by CGPA descending using Comparator + lambda
        list.sort((a, b) -> Double.compare(b.getCgpa(), a.getCgpa()));
        System.out.println(list);
    }
}
\end{lstlisting}

\section{Demo / Observation Task}
Use this block as a short reproducible demo or observation task.
\begin{itemize}
  \item Reproduce the lecture's main example around \textit{Core interfaces: \texttt{Iterable}, \texttt{Collection}, \texttt{Iterator}} once without looking at the solution first.
  \item Change one input, rule, or design choice in Campus Resource Manager and predict the result before checking it.
  \item Record one success case, one edge case, and one failure case so the idea becomes observable instead of purely theoretical.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Practice Check (with brief solutions)}
\begin{enumerate}
  \item Why do we use \texttt{Iterator.remove()} instead of \texttt{list.remove(...)} inside a for-each loop?\par
  \textbf{Answer:} to avoid concurrent modification issues; iterator removal is safe during traversal.

  \item Which list is better for random access by index?\par
  \textbf{Answer:} \texttt{ArrayList}.

  \item What is the difference between \texttt{Comparable} and \texttt{Comparator}?\par
  \textbf{Answer:} \texttt{Comparable} defines natural order inside class; \texttt{Comparator} defines external/custom order.
\end{enumerate}

\section{Self-Study Practice (no solutions)}
\begin{enumerate}
  \item Create a list of integers and remove all odd numbers using an iterator.
  \item Sort a list of strings by length using a comparator lambda.
  \item Implement \texttt{Comparable} in a class \texttt{Book} to sort by title.
\end{enumerate}

\section{Further Study}
\begin{itemize}
  \item Revisit the prerequisite references if any recall bullet still feels weak.
  \item Rework the lecture example with one changed assumption, input, or design choice.
  \item Read the textbook and documentation references listed in the source line for a second explanation of the same topic.
  \item Apply the same idea once inside Campus Resource Manager so that the concept moves from theory to design practice.
\end{itemize}
\notesources{\OOPSUnitVSources}

\section{Exit Questions (with sample answers)}
\textbf{Q1.} Why does Java provide a Collections Framework?\par
\textbf{Sample answer:} It provides standard, reusable data structures and algorithms, reducing bugs and improving productivity compared to writing custom structures every time.

\vspace{0.6em}
\textbf{Q2.} When should you prefer \texttt{Comparator} over \texttt{Comparable}?\par
\textbf{Sample answer:} Prefer \texttt{Comparator} when you need multiple sorting orders or when you cannot/should not modify the class to define natural ordering.

\vspace{0.6em}
\textbf{Q3.} What decides between ArrayList and LinkedList in practice?\par
\textbf{Sample answer:} Access patterns, update patterns, readability, and real measured needs decide the choice.

\end{document}
